---
title: "Analysis"
author: "Kaija Gahm"
date: "1/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load libraries
```{r, echo = F}
source("libraries.R")
```

# Load data
```{r}
load("data/outputs/dat3.Rda")
load("data/outputs/wild.Rda")

# without NA's
load("data/outputs/wild_nonas.Rda")
load("data/outputs/dat3_nonas.Rda")
load("data/outputs/wild_nonas_std.Rda")
load("data/outputs/dat3_nonas_std.Rda")
```

# Summary stats
## Number of individuals
```{r}
length(unique(wild_nonas$tadpole)) + length(unique(dat3_nonas$tadpole))

dat3_nonas_forsummary <- dat3_nonas %>%
  select(-c(trial, sline_speed_mm_s, T_initial, camera)) %>%
  distinct() %>%
  mutate(clutch_only = str_extract(clutch, "(?<= )[[:digit:]]{1}"))

# Low treatment
dat3_nonas_forsummary %>%
  filter(treatment == "Low") %$% 
  table(pond, clutch_only)

# High treatment
dat3_nonas_forsummary %>%
  filter(treatment == "High") %$%
  table(pond, clutch_only)

clutch_n <- dat3_nonas_forsummary %>%
  group_by(treatment, clutch) %>%
  summarize(n = n())

clutch_n %>% summarize(median = median(n),
                       max = max(n),
                       min = min(n))

# Lab-reared by pond
pond_n <- dat3_nonas_forsummary %>%
  group_by(treatment, pond) %>%
  summarize(n = n())
pond_n


# Wild tadpoles
wild_pond_n <- wild_nonas %>% 
  select(-c(trial, sline_speed_mm_s, T_initial, camera)) %>%
  distinct() %$% 
  table(pond)
median(wild_pond_n)
max(wild_pond_n)
min(wild_pond_n)
```


# Load consensus model
```{r}
load("data/outputs/consensus_mod.Rda")
summary(consensus_mod)
```

# Variance partitioning
### Manual calculation of residuals
#### Retaining the individual speeds and comparing them to different things
```{r}
overall <- dat3 %>% select(treatment, pond, clutch, tadpole, trial, sline_speed_mm_s) %>% 
  group_by(treatment) %>%
  mutate(mean = mean(sline_speed_mm_s),
                resid = sline_speed_mm_s - mean) %>%
  mutate(level = "overall")

indiv <- overall %>% group_by(treatment, tadpole) %>% 
  mutate(mean = mean(sline_speed_mm_s),
         resid = sline_speed_mm_s - mean) %>%
  mutate(level = "indiv")

clutch <- overall %>% group_by(treatment, clutch) %>%
  mutate(mean = mean(sline_speed_mm_s),
         resid = sline_speed_mm_s - mean) %>%
  mutate(level = "clutch")

pond <- overall %>% group_by(treatment, pond) %>%
  mutate(mean = mean(sline_speed_mm_s),
         resid = sline_speed_mm_s - mean) %>%
  mutate(level = "pond")

wild_overall <- wild %>% select(treatment, pond, tadpole, trial, sline_speed_mm_s) %>%
  group_by(treatment) %>%
  mutate(mean = mean(sline_speed_mm_s),
         resid = sline_speed_mm_s - mean) %>%
  mutate(level = "overall")

wild_indiv <- wild_overall %>% group_by(treatment, tadpole) %>%
  mutate(mean = mean(sline_speed_mm_s),
         resid = sline_speed_mm_s - mean) %>%
  mutate(level = "indiv")

wild_pond <- wild_overall %>% group_by(treatment, pond) %>%
  mutate(mean = mean(sline_speed_mm_s),
         resid = sline_speed_mm_s - mean) %>%
  mutate(level = "pond")

resids <- rbind(overall, indiv, clutch, pond, wild_overall, wild_indiv, wild_pond) # put them all together

resids$level <- factor(resids$level, levels = c("overall", "pond", "clutch", "indiv"))

# Density plot
resids %>% ggplot(aes(x = resid, col = treatment, fill = treatment))+
  geom_density(alpha = 0.4, adjust = 2)+
  facet_grid(rows = vars(level), cols = vars(treatment))

# Ridge plot
resids %>% ggplot(aes(x = resid, fill = level))+
  geom_density_ridges(aes(y = level),
                      scale = 3)+
                      #bandwidth = 0.006)+
  facet_wrap(~treatment)+
  scale_y_discrete(expand = expand_scale(mult = c(0.01, 0.8)))+
  theme(legend.position = "none")

# get just the variance of the resids
var_resids <- resids %>% 
  group_by(treatment, level) %>%
  summarize(var = var(resid)) %>% ungroup

var_resids <- rbind(var_resids, c("Wild", "clutch", NA)) %>% mutate(var = as.numeric(var))

# Line plot
var_resids %>% ggplot(aes(x = level, y = var, col = treatment, group = treatment))+
  geom_line()+
  geom_point(data = var_resids %>% filter(treatment == "Wild", level == "indiv"))
``` 

#### Retaining the individual speeds and comparing them to different things
```{r}
overallsmall <- overall %>% select(treatment, mean, resid, level)

indivsmall <- indiv %>% select(treatment, mean, resid, level)

clutch <- indiv %>% 
  select(treatment, pond, clutch, tadpole, mean) %>% 
  distinct() %>% # get rid of multiple trials
  rename(indiv_mean = mean) %>%
  group_by(treatment, clutch) %>%
  mutate(clutch_mean = mean(indiv_mean),
         resid = indiv_mean - clutch_mean) %>%
  select(-c(indiv_mean, tadpole)) %>%
  rename(mean = clutch_mean) %>% ungroup() %>%
  mutate(level = "clutch")
clutchsmall <- clutch %>% select(treatment, mean, resid, level)

pond <- clutch %>% select(-resid) %>%
  distinct() %>%
  rename(clutch_mean = mean) %>%
  group_by(treatment, pond) %>%
  mutate(pond_mean = mean(clutch_mean),
         resid = clutch_mean - pond_mean) %>%
  select(-c(clutch_mean, clutch)) %>%
  rename(mean = pond_mean) %>% ungroup() %>%
  mutate(level = "pond")
pondsmall <- pond %>% select(treatment, mean, resid, level)

wild_overallsmall <- wild_overall %>% select(treatment, mean, resid, level)

wild_indiv <- wild_overall %>% select(treatment, pond, tadpole, sline_speed_mm_s) %>%
  group_by(tadpole) %>% 
  mutate(indiv_mean = mean(sline_speed_mm_s),
         resid = sline_speed_mm_s - indiv_mean) %>%
  rename(mean = indiv_mean) %>% ungroup() %>%
  mutate(level = "indiv")
wild_indivsmall <- wild_indiv %>% select(treatment, mean, resid, level)

wild_pond <- wild_indiv %>% select(-resid) %>%
  distinct() %>%
  rename(indiv_mean = mean) %>%
  group_by(treatment, pond) %>%
  mutate(pond_mean = mean(indiv_mean),
         resid = indiv_mean - pond_mean) %>%
  select(-c(indiv_mean)) %>%
  rename(mean = pond_mean) %>% ungroup() %>%
  mutate(level = "pond")
wild_pondsmall <- wild_pond %>% select(treatment, mean, resid, level)

resids_cumulative <- bind_rows(overallsmall, indivsmall, clutchsmall, pondsmall, wild_overallsmall, wild_indivsmall, wild_pondsmall)
resids_cumulative$level <- factor(resids_cumulative$level, levels = c("overall", "pond", "clutch", "indiv"))
save(resids_cumulative, file = "data/outputs/resids_cumulative.Rda")

# Density plot
resids_cumulative %>% ggplot(aes(x = resid, col = treatment, fill = treatment))+
  geom_density(alpha = 0.4, adjust = 2)+
  facet_grid(rows = vars(level), cols = vars(treatment))

# Ridge plot
resids_cumulative %>% ggplot(aes(x = resid, fill = level))+
  geom_density_ridges(aes(y = level),
                      scale = 3,
                      bandwidth = 0.006)+
  facet_wrap(~treatment)+
  scale_y_discrete(expand = expand_scale(mult = c(0.01, 0.8)))+
  theme(legend.position = "none")+
  theme_minimal()+
  xlab("Residual")+
  ylab("") +
  guides(fill = guide_legend(title = "Level of \n hierarchy"))

# get just the variance of the resids
var_resids_cumulative <- resids_cumulative %>% 
  group_by(treatment, level) %>%
  summarize(var = var(resid)) %>% ungroup

var_resids_cumulative <- rbind(var_resids_cumulative, c("Wild", "clutch", NA)) %>% mutate(var = as.numeric(var))

# Line plot
var_resids_cumulative %>% ggplot(aes(x = level, y = var, col = treatment, group = treatment))+
  geom_line()+
  geom_point(data = var_resids_cumulative %>% filter(treatment == "Wild", level == "indiv"))
```

```{r}
load("data/outputs/consensus_mod.Rda")
sds <- summary(consensus_mod)$varcor %>% as.data.frame()

indiv_variance <- sds$vcov[sds$grp == "indiv:clutch"]
clutch_variance <- sds$vcov[sds$grp == "clutch"]
resid_variance <- sds$vcov[sds$grp == "Residual"]

tot_rand_variance <- indiv_variance + clutch_variance + resid_variance
iccc_indiv <- indiv_variance/tot_rand_variance
iccc_clutch <- clutch_variance/tot_rand_variance

iccc_indiv # individual-level variance is a much larger component than clutch-level variance.
iccc_clutch
```

# Plasticity
## Differences in means
### Individual means
```{r}
indiv_means <- dat3_nonas %>% group_by(indiv) %>% 
  summarize(indiv_mean_speed = mean(sline_speed_mm_s),
            clutch = clutch[1],
            treatment = treatment[1],
            pond = pond[1]) 

head(indiv_means)
```

### Clutch means
```{r}
clutch_means <- indiv_means %>% 
  group_by(clutch, treatment) %>% 
  summarize(clutch_mean_speed = mean(indiv_mean_speed))

clutch_means %>% arrange(clutch)
clutch_means %$% table(clutch)
```

#### Remove clutches not in both treatments
```{r}
clutch_means <- clutch_means %>% 
  group_by(clutch) %>%
  mutate(freq = n()) %>%
  ungroup() %>%
  filter(freq == 2) %>%
  select(-freq)
```

###Treatment differences
```{r}
treatment_diffs <- clutch_means %>%
  group_by(clutch) %>%
  summarize(diff = clutch_mean_speed[treatment == "Low"] - clutch_mean_speed[treatment == "High"])

treatment_diffs %>% ggplot(aes(x = diff))+
  geom_histogram()
```
What I've calculated here is D, the "scope of plastic response" (Stearns 1992, cited in Valladares _et al._ 2006): mean at low resource availability - mean at high resource availability. (note low vs. high is switched from the paper because in this case high temp ~ slower speed.)

## Individuals vs. means
### Opposite treatment means
```{r}
clutch_means <- clutch_means %>%
  group_by(clutch) %>%
  mutate(opposite_clutch_mean = case_when(treatment == "High" ~ clutch_mean_speed[treatment == "Low"],
                                   treatment == "Low" ~ clutch_mean_speed[treatment == "High"]))

head(clutch_means)
```

### Join to individual means
```{r}
indiv_means <- indiv_means %>%
  left_join(clutch_means, by = c("treatment", "clutch"))

head(indiv_means)
```

### Indiv/opp clutch differences
If we always subtract individual - mean, or always mean - individual, the distribution will be centered around 0 because we'll be subtracting a mixture of high-low and low-high. Need to be consistent with temperatures, which means inconsistent with direction of subtraction.

We're going to default to always subtracting low - high. 
```{r}
indiv_means <- indiv_means %>%
  mutate(indiv_opp_clutch_diff = case_when(treatment == "High" ~ opposite_clutch_mean - indiv_mean_speed,
                                           treatment == "Low" ~ indiv_mean_speed - opposite_clutch_mean))
```

#### Simplify data
```{r}
plasticity <- indiv_means %>% select(-c(clutch_mean_speed)) %>% rename(D = indiv_opp_clutch_diff)
```

#### Fit model predicting difference by pond
```{r}
head(plasticity)
plasticity <- plasticity %>% mutate(clutch_only = factor(str_extract(clutch, pattern = "(?<= )[[:digit:]](?=$)")))
save(plasticity, file = "data/outputs/plasticity.Rda")

# Plasticity model: predict D by pond and clutch
pls_mod <- lmer(D ~ pond + (1|clutch), data = plasticity)
summary(pls_mod)

plot(pls_mod) # plot residuals: they look good.

# Plot effect sizes
sjPlot::plot_model(pls_mod, type = "eff", title = "")
# SE's are computed after taking away the variance by clutch. 

# Plot all D points
plasticity %>% ggplot(aes(x = pond, y = D))+
  geom_violin()+
  geom_jitter(aes(col = clutch_only), width = 0.1)+
  scale_color_viridis(discrete = T) # no patterns by pond _or_ clutch within pond. Yikes!

# Get order of ponds by mean plasticity
mean_pond_order <- plasticity %>% 
  group_by(pond) %>% 
  summarize(mean = mean(D, na.rm = T)) %>%
  arrange(-mean) %>%
  pull(pond) %>%
  as.character()

# Make plot
## Define helper functions
reorder_within <- function(x, by, within, fun = mean, sep = "___", ...) {
  new_x <- paste(x, within, sep = sep)
  stats::reorder(new_x, by, FUN = fun)
}

scale_x_reordered <- function(..., sep = "___") {
  reg <- paste0(sep, ".+$")
  ggplot2::scale_x_discrete(labels = function(x) gsub(reg, "", x), ...)
}

## make the plot
plasticity %>%
  mutate(pond = fct_relevel(pond, mean_pond_order)) %>%
  filter(!is.na(clutch_only)) %>%
  ggplot(aes(x = reorder_within(clutch_only, -D, pond, mean), y = D))+
  geom_rect(aes(fill = pond),
            xmin = -Inf, xmax = Inf,
            ymin = -Inf, ymax = Inf, 
            alpha = 0.3)+
  scale_fill_manual(values = rep(c("lightskyblue1", "skyblue2"), 5))+
  geom_boxplot()+
  scale_x_reordered()+
  facet_grid(cols = vars(pond), scales = "free_x")+
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    legend.position = "none"
  )+
  ylab("Scope of plastic response (low temp - high temp)")+
  xlab("Pond")
```

### Clutch mean differences
```{r}
clutch_means
clutch_mean_diffs <- clutch_means %>% 
  group_by(clutch) %>%
  filter(treatment == "Low") %>%
  summarize(D = clutch_mean_speed - opposite_clutch_mean) %>% # subtract low - high
  mutate(pond = str_extract(clutch, "(?<=^)[[:alnum:]]+(?= )"))

head(clutch_mean_diffs)
```

#### Fit model predicting difference by pond
```{r}
# Plasticity model: predict D by pond and clutch
pls_mod_means <- lm(D ~ pond, data = clutch_mean_diffs)
summary(pls_mod_means)

plot(pls_mod_means) # plot residuals: they look good.

# Plot effect sizes
sjPlot::plot_model(pls_mod_means, type = "eff", title = "Predicted values of D, means")
sjPlot::plot_model(pls_mod, type = "eff", title = "Predicted values of D, individuals")

# Plot all D points
clutch_mean_diffs %>% ggplot(aes(x = pond, y = D))+
  geom_violin()+
  geom_jitter(width = 0.1)
```


# Morphology vs. Burst Speed
Does morphology help to explain burst speed? And does this relationship hold in wild populations?

## Among wild tadpoles...
### Explore wild data
```{r}
# Any differences by pond?
wild_nonas %>% ggplot(aes(x = mass, y = sline_speed_mm_s, col = T_initial))+
  geom_smooth(method = "lm", col = "gray", se = F)+
  geom_point()+
  facet_wrap(~pond)+
  scale_color_viridis()

# Size
wild_nonas %>% 
  ggplot(aes(x = size, y = sline_speed_mm_s))+
  geom_smooth(aes(col = pond), method = "lm", se = F)+
  geom_smooth(method = "lm", se = F, col = "darkgrey", lty = 2)+
  geom_point(aes(col = pond))+
  scale_color_viridis(discrete = TRUE)

# Gosner stage
wild_nonas %>% 
  ggplot(aes(x = gs, y = sline_speed_mm_s))+
  geom_smooth(aes(col = pond), method = "lm", se = F)+
  geom_smooth(method = "lm", se = F, col = "darkgrey", lty = 2)+
  geom_point(aes(col = pond))+
  scale_color_viridis(discrete = TRUE)

# Mass
wild_nonas %>% 
  ggplot(aes(x = mass, y = sline_speed_mm_s))+
  geom_smooth(aes(col = pond), method = "lm", se = F)+
  geom_smooth(method = "lm", se = F, col = "darkgrey", lty = 2)+
  geom_point(aes(col = pond))+
  scale_color_viridis(discrete = TRUE)

# Shape PC1
wild_nonas %>% 
  ggplot(aes(x = PC1, y = sline_speed_mm_s))+
  geom_smooth(aes(col = pond), method = "lm", se = F)+
  geom_smooth(method = "lm", se = F, col = "darkgrey", lty = 2)+
  geom_point(aes(col = pond))+
  scale_color_viridis(discrete = TRUE)

# Shape PC2
wild_nonas %>% 
  ggplot(aes(x = PC2, y = sline_speed_mm_s))+
  geom_smooth(aes(col = pond), method = "lm", se = F)+
  geom_smooth(method = "lm", se = F, col = "darkgrey", lty = 2)+
  geom_point(aes(col = pond))+
  scale_color_viridis(discrete = TRUE)

# Initial water temp
wild_nonas %>%
  ggplot(aes(x = T_initial, y = sline_speed_mm_s))+
  geom_smooth(aes(col = pond), method = "lm", se = F)+
  geom_smooth(method = "lm", se = F, col = "darkgrey", lty = 2)+
  geom_point(aes(col = pond))+
  scale_color_viridis(discrete = TRUE)
```

# Comparison between lab-reared and wild for the same pond. 
## Combine datasets
```{r}
names(wild_nonas)
names(dat3_nonas)

all <- bind_rows(wild_nonas, dat3_nonas)

indiv_means <- all %>% 
  group_by(pond, treatment, tadpole) %>%
  summarize(speed = mean(sline_speed_mm_s),
            mass = mass[1])
head(indiv_means)
save(indiv_means, file = "data/outputs/indiv_means.Rda")
```

## Plot
```{r}
indiv_means %>%
  ggplot(aes(x = mass, y = speed, col = treatment))+
  geom_smooth(method = "lm")+
  facet_wrap(~pond)+
  theme_minimal()
```

## Model predictions
```{r}
# Load data
load("data/outputs/dat3_nonas.Rda")
load("data/outputs/wild_nonas.Rda")

load("data/outputs/dat3_nonas_std.Rda")
load("data/outputs/wild_nonas_std.Rda")

# Load models
load("data/outputs/consensus_mod.Rda")
load("data/outputs/replica_mod_wild.Rda")
```

### Development and allocation
```{r}
plot.alloc <- dat3_nonas %>% ggplot(aes(x = gs, y = mass, col = treatment))+ # plot lab-reared tadpoles
  geom_point()+
  geom_smooth(method = "lm")+
  geom_point(data = wild_nonas, color = "black")+ # add points and line for wild tadpoles
  geom_smooth(data = wild_nonas, method = "lm", col = "black")+
  scale_color_manual(values = c("orange", "dodgerblue3"))
plot.alloc
```
It looks like there is no real difference in allocation between the lab treatments, but wild tadpoles are generally larger at all stages. Worth noting that the wild tads have much more variance. Slopes (allocation rate) is pretty similar. -Andis

### Partial regressions
Display the effect of a focal variable while holding the other variables at their averages (or medians, for Gosner stage). In this case, we're holding other variables at the mean _of the lab tadpoles_

For these partial regressions, we'll work with _unstandardized_ data.

#### Mean-constant data (GS)
Hold all variables constant at their mean values for lab tadpoles except for Gosner stage. Mass_at_stage will be held to 0, since those values are residuals.
```{r}
# Load data
load("data/outputs/dat3_nonas.Rda")
load("data/outputs/wild_nonas.Rda")

# Load models
load("data/outputs/consensus_mod.Rda")
load("data/outputs/consensus_mod_wild.Rda")

# Make new datasets
newlab.gs <- dat3_nonas %>% # make new means/medians dataset (lab)
  select(tadpole, clutch, mass_at_stage, treatment, pond, T_initial, gs, PC2) %>%
  mutate(T_initial = mean(T_initial),
         mass_at_stage = mean(mass_at_stage),
         PC2 = mean(PC2))

newwld.gs <- wild_nonas %>%
  select(tadpole, mass_at_stage, avg_pond_temp, pond, T_initial, gs, PC2) %>%
  mutate(T_initial = mean(T_initial),
         mass_at_stage = mean(mass_at_stage),
         avg_pond_temp = mean(avg_pond_temp),
         PC2 = mean(PC2))

# Add predictions
newlab.gs <- newlab.gs %>%
  mutate(fit.nore = predict(consensus_mod, newdata = newlab.gs, re.form = ~0),
         fit.re = predict(consensus_mod, newdata = newlab.gs, re.form = NULL))

newwld.gs <- newwld.gs %>%
  mutate(fit.nore = predict(consensus_mod_wild, newdata = newwld.gs, re.form = ~0),
         fit.re = predict(consensus_mod_wild, newdata = newwld.gs, re.form = NULL))
```

#### Plot (GS)
```{r}
plot.gs <- newlab.gs %>%
  ggplot(aes(x = gs, y = fit.re, col = treatment))+
  geom_point()+ # lab points
  geom_point(data = newwld.gs, col = "black")+
  geom_smooth(aes(y = fit.nore), 
              method = "lm", se = F)+
  geom_smooth(data = newwld.gs, aes(y = fit.nore),
              method = "lm", se = F, col = "black")+
  scale_color_manual(values = c("orange", "dodgerblue3"))+
  ylab("Predicted speed (mm/s)")+
  xlab("Gosner stage")+
  theme_minimal()
plot.gs
```

#### Mean-constant data (resid)
```{r}
# Make new datasets
newlab.sas <- dat3_nonas %>%
  select(tadpole, clutch, mass_at_stage, treatment, pond, T_initial, gs, PC2) %>%
  mutate(T_initial = mean(T_initial),
         gs = median(gs),
         PC2 = mean(PC2))

newwld.sas <- wild_nonas %>%
  select(tadpole, mass_at_stage, avg_pond_temp, pond, T_initial, gs, PC2) %>%
  mutate(T_initial = mean(T_initial),
         gs = median(gs),
         avg_pond_temp = mean(avg_pond_temp),
         PC2 = mean(PC2))

# Add predictions
newlab.sas <- newlab.sas %>%
  mutate(fit.nore = predict(consensus_mod, newdata = newlab.sas, re.form = ~0),
         fit.re = predict(consensus_mod, newdata = newlab.sas, re.form = NULL))

newwld.sas <- newwld.sas %>%
  mutate(fit.nore = predict(consensus_mod_wild, newdata = newwld.sas, re.form = ~0),
         fit.re = predict(consensus_mod_wild, newdata = newwld.sas, re.form = NULL))
```

#### Save datasets
```{r}
save(newlab.gs, file = "data/outputs/newlab.gs.Rda")
save(newwld.gs, file = "data/outputs/newwld.gs.Rda")
save(newlab.sas, file = "data/outputs/newlab.sas.Rda")
save(newwld.sas, file = "data/outputs/newwld.sas.Rda")
```

#### Plot (mass at stage)
```{r}
plot.sas <- newlab.sas %>%
  ggplot(aes(x = mass_at_stage, y = fit.re, col = treatment)) +
  geom_point() +
  geom_point(data = newwld.sas, col = "black") +
  geom_smooth(data = newlab.sas, aes(y = fit.nore), 
              method = 'lm', 
              se = F) +
  geom_smooth(data = newwld.sas, aes(y = fit.nore), 
              method = 'lm', 
              se = F, col = "black") +
  scale_color_manual(values = c("orange", "dodgerblue3"))+
  ylab("Predicted speed (mm/s)")+
  xlab("Mass at stage")+
  theme_minimal()
plot.sas
```

#### Plot (both)
```{r}
cowplot::plot_grid(plot.gs, plot.sas)
```

### Without points (RE by pond)
I think the plots above are a little misleading because they don't show clear how much variaition there is in the wild compared to the lab tadpoles. The lab animals have really tight distributions, but the wild animals are all over the place and vary a ton just due to the pond that they were from.

Below I've dropped the points and instead am plotting the random effects by pond to show the variation. -Andis
#### GS
```{r}
plot.gs.re <- newlab.gs %>%
  ggplot(aes(x = gs, y = fit.re, col = treatment)) +
  geom_smooth(aes(y = fit.re, group = interaction(pond, treatment)), 
              method = 'lm', se = F, size = 0.5) +
  geom_smooth(aes(y = fit.nore), 
              method = 'lm', se = F, size = 2) +
  geom_smooth(data = newwld.gs, aes(y = fit.re, group = pond), 
              method = 'lm', se = F, col = 1, size = 0.5) +
  geom_smooth(data = newwld.gs, aes(y = fit.nore), 
              method = 'lm', se = F, col = 1, size = 2) +
  scale_color_manual(values = c("orange", "dodgerblue3"))+
  ylab("Predicted speed (mm/s)")+
  xlab("Gosner stage")+
  theme_minimal()
plot.gs.re
```

#### Mass at stage
```{r}
plot.sas.re <- newlab.sas %>% ggplot(aes(x = mass_at_stage, y = fit.re, col = treatment)) +
  geom_smooth(aes(y = fit.re, group = interaction(pond, treatment)), 
              method = 'lm', se = F, size = 0.5) +
  geom_smooth(aes(y = fit.nore), 
              method = 'lm', se = F, size = 2) +
  geom_smooth(data = newwld.sas, aes(y = fit.re, group = pond), 
              method = 'lm', se = F, col = 1, size = 0.5) +
  geom_smooth(data = newwld.sas, aes(y = fit.nore), 
              method = 'lm', se = F, col = 1, size = 2) +
  scale_color_manual(values = c("orange", "dodgerblue3")) +
  ylab("Predicted speed (mm/s)")+
  xlab("Mass at stage")+
  theme_minimal()
plot.sas.re
```

#### Both
```{r}
cowplot::plot_grid(plot.gs.re, plot.sas.re)
```

#### All four
```{r}
cowplot::plot_grid(plot.gs, plot.sas, plot.gs.re, plot.sas.re, nrow = 2)
```


