---
title: "Analysis"
author: "Kaija Gahm"
date: "1/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load libraries
```{r, echo = F}
source("libraries.R")
```

# Load data
## Select relevant columns and change data types
```{r}
load("data/outputs/stats_fast3.Rda")
dat3 <- stats_fast3 %>% select(tadpole, trial, indiv, treatment, sline_speed_m_s, PC1, PC2, pond, size, clutch, mass, gs, T_initial) %>% ungroup() %>% mutate(treatment = factor(treatment),
                                                      pond = factor(pond),
                                                      clutch = factor(clutch)) %>% 
  filter(treatment %in% c("High", "Low"), !is.na(PC1)) %>% droplevels()
```

## Remove observations that are NA for mass
```{r}
dat3 <- dat3 %>% filter(!is.na(mass))
```

# Code random effects explicitly
We should have used different letters for the individuals in the two temperature treatments, because strictly speaking we should have labeled the individuals before choosing which ones to put into which temperature treatment. We did select those randomly, but the naming scheme that we chose does not reflect that.

We don't want to change the letters on the tadpoles because that would make the data be out of sync with all of the labels on everything up to this point. Instead, we're going to stick with our scheme of appending "high" or "low" to the names to distinguish e.g. CPS_01_A in high/low treatments (since that is NOT the same individual.)
```{r}
dat3 <- dat3 %>% # remove tadpole column
  mutate(clutch = factor(paste(pond, clutch)), # code clutches explicitly
         indiv = factor(paste(clutch, indiv, treatment))) # code individuals explicitly
locate.nas(dat3)
```


Hierarchical mixed model, incorporating pond within clutch and accounting for multiple trials per individual
Dat3: fastest 3 trials per tadpole
```{r}
mod1 <- lmer(sline_speed_m_s ~ treatment * mass + (1|pond/clutch) + (1|indiv), data = dat3)
summary(mod1)

dat3$predicted_speed <- predict(mod1, re.form = NA, newdata = dat3)
dat3$predicted_speed_withresids <- predict(mod1, re.form = NA, newdata = dat3) + resid(mod1, newdata = dat3)

# Compare ggplots
dat3 %>% ggplot(aes(x = mass, y = sline_speed_m_s, col = treatment))+
  geom_point()+
  geom_smooth(method = "lm")+
  ylab("speed")+
  ggtitle("Original data")

dat3 %>% ggplot(aes(x = mass, y = predicted_speed, col = treatment))+
  geom_point()+
  geom_smooth(method = "lm")+
  ylab("predicted_speed")+
  ggtitle("Predicted data")

dat3 %>% ggplot(aes(x = mass, y = predicted_speed_withresids, col = treatment))+
  geom_point()+
  geom_smooth(method = "lm")+
  ylab("predicted_speed_withresids")+
  ggtitle("Predicted data with residuals")
```
The similarity between these graphs shows us that we have correctly specified the structure of our model.


