---
title: "Analysis"
author: "Kaija Gahm"
date: "11/21/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup
## Load libraries
```{r, echo = FALSE}
source("libraries.R")
source("process_coordinates_funs.R")
```

## Load data
```{r}
load("data/outputs/bursts_allcrops.Rda")
load("data/outputs/allmorpho.Rda")
load("data/outputs/scoords_allcrops.Rda")
```

## Summary stats
### Calculate stats
```{r}
burst_stats <- bursts_allcrops %>% 
  group_by(tadpole, trial, crop) %>%
  summarize(avgspeed = mean(speed, na.rm = T),
            maxspeed = max(speed, na.rm = T),
            maxspeedma5 = max(speedma5, na.rm = T)) %>%
  as.data.frame()

burst_info <- bursts_allcrops %>% select(tadpole, trial, pond, clutch, indiv, treatment, year, month, day, camera, crop) %>% distinct()

burst_stats <- left_join(burst_stats, burst_info, by = c("tadpole", "trial", "crop"))
```

### Add straight line distance from coords
```{r}
coords_stats <- scoords_allcrops %>%
  group_by(tadpole, trial, crop) %>%
  filter(row_number() %in% c(1, n())) %>% 
  summarize(xinit = x[1],
            xfin = x[n()], # note that I use x[n()] instead of x[2] here because there was one burst that didn't have a second row, since it was exactly 31 frames long. This is super rare, but it's worth making the code robust so the second half second speed will show up as 0 and not as NA.
            yinit = y[1],
            yfin = y[n()],
            xdiff = xfin-xinit,
            ydiff = yfin-yinit,
            slinedist = sqrt(xdiff^2 + ydiff^2),
            time = frame_burst[n()]/60,
            sline_speed_m_s = slinedist/time) %>% select(tadpole, trial, crop, slinedist, time, sline_speed_m_s)

burst_stats <- left_join(burst_stats, coords_stats, by = c("tadpole", "trial", "crop"))

# Sanity check on speeds: We're seeing speeds ranging from 0 to 15 centimeters/sec, which makes sense.
burst_stats %>% ggplot(aes(x = sline_speed_m_s, col = crop, fill = crop))+
  geom_density(alpha = 0.4)
```

### Save stats
```{r}
#save(burst_stats, file = "data/outputs/burst_stats.Rda")
```

## Which time crop to use
"Although a full second was measured for all tadpoles, most showed signs of slowing during the second half second (paired t‐tests comparing first and second half seconds: Bufo: t202 = 12 Scaphiopus t202 = 21, P < 0·001 for both)" (Arendt 2003, Functional Ecology)
### Comparing avgspeed for first and second half seconds
```{r}
test <- burst_stats %>% filter(crop %in% c("First half second", "Second half second"))

# Paired t-test for first and second half seconds, excluding bursts that don't have both.
problems <- test %>% group_by(tadpole, trial) %>% summarize(freq = n()) %>% filter(freq < 2) # identify problems

# Remove problems
test <- test %>% filter(!(paste(tadpole, trial) %in% paste(problems$tadpole, problems$trial))) %>% mutate(unique = paste(tadpole, trial)) %>% select(unique, crop, avgspeed, sline_speed_m_s)

# Run paired t-test
t.test(avgspeed ~ crop, paired = T, data = test) # I'm confused about which direction this is going.

# Sanity check
test %>% ggplot(aes(x = avgspeed, col = crop, fill = crop))+
  geom_density(alpha = 0.4)
```
Interestingly, the t-test does not reveal a significant difference in means. That's odd. 

### Comparing straight line speed for first and second half seconds
```{r}
# Run paired t-test
t.test(sline_speed_m_s ~ crop, data = test) # alternative hypothesis that second half sec has a smaller mean than first half sec

# Sanity check
test %>% ggplot(aes(x = sline_speed_m_s, col = crop, fill = crop))+
  geom_density(alpha = 0.4)
```
We do see a difference in straight line speed between the first and second half sec


## First half sec average across trials
```{r}
# By trial run (tadpole x date)
burst_stats_avg_tadpole <- burst_stats %>% filter(crop == "First half second") %>% group_by(tadpole) %>%
  summarize(avgavgspeed = mean(avgspeed),
            avgmaxspeed = mean(maxspeed),
            avgmaxspeedma5 = mean(maxspeedma5),
            pond = pond[1],
            clutch = clutch[1],
            indiv = indiv[1],
            treatment = treatment[1],
            year = year[1],
            month = month[1],
            day = day[1],
            camera = camera[1],
            avgslinedist = mean(slinedist),
            time = time[1],
            avgslinespeed_ms = mean(sline_speed_m_s)) %>%
  as.data.frame()
```

## Join stats to morpho data
```{r}
# Cleaning
#First, fix that one space.
burst_stats$tadpole[burst_stats$tadpole == "LO_W_09 2_Wild"] <- "LO_W_09_2_Wild"

# Join stats to morpho data
stats <- burst_stats %>% select(-c(pond, clutch)) %>% 
  left_join(allmorpho, by = c("tadpole", "treatment")) %>%
         rename(burst_year = "year",
         burst_month = "month",
         burst_day = "day",
         photo_date = "date") %>% select(-c(indiv.y, id, treat)) %>% rename(indiv = indiv.x)

locate.nas(stats) # weirdness.
problems <- stats %>% filter(is.na(PC1)) %>% select(tadpole, trial, crop, treatment) %>% distinct()
problemtadpoles <- unique(problems$tadpole)
problemtadpoles
problemtadpoles %in% allmorpho$tadpole # we just don't have morpho data for these tadpoles; it's not a faulty join.
problemtadpoles
```

## Exploratory plotting
```{r}
tadpoles <- unique(bursts_allcrops$tadpole)

bursts_allcrops %>% filter(tadpole == tadpoles[100], crop == "First half second") %>% 
  ggplot(aes(x = frame_burst, y = speedma5, col = factor(trial)))+
  geom_line()+
  xlab("Frame")+
  ylab("Speed (5-frame moving average)")
```

```{r}
head(stats)

# Straight line speed by pond (averaging across trials for each tadpole)
burst_stats_avg_tadpole %>% ggplot()+
  geom_violin(aes(x = factor(pond), y = avgslinespeed_ms), fill = "lightblue")+
  ggtitle("Burst speed by pond of origin")+
  xlab("Pond")+
  ylab("Burst speed (m/s)") + theme_minimal()

# Straight line speed by treatment (averaging across trials for each tadpole)
burst_stats_avg_tadpole %>% ggplot()+
  geom_boxplot(aes(x = treatment, y = avgslinespeed_ms, fill = treatment))+
  ggtitle("Burst speed by treatment")+
  xlab("Treatment")+
  ylab("Speed (m/s)")+
  labs(fill = "Temperature \n treatment")

burst_stats_avg_tadpole %>% ggplot()+
  geom_violin(aes(x = treatment, y = avgslinespeed_ms, fill = treatment))+
  ggtitle("Burst speed by treatment")+
  xlab("Treatment")+
  ylab("Speed (m/s)")+
  labs(fill = "Temperature \n treatment")+ theme_minimal()
```


### Does speed decrease by trial number?
```{r}
stats %>% filter(crop == "First half second", trial < 5, sline_speed_m_s < 0.2) %>%
  ggplot(aes(x = jitter(trial), y = sline_speed_m_s, col = treatment))+
  geom_point()+
  geom_smooth(method = 'lm', formula = y ~ x)+
  ggtitle("Straight line speed by trial number")+
  ylab("Straight line speed (first half second)")+
  xlab("Trial number")
```
```{r}
# Model the decrease in speed
tiring_model <- lm(sline_speed_m_s ~ trial + treatment, data = stats)
summary(tiring_model)

# There is an effect, but the effect is definitely small.

# Limit it to under 5 trials
tiring_model_under5 <- lm(sline_speed_m_s ~ trial + treatment, data = stats[stats$trial < 5,])
summary(tiring_model_under5)

# Less of an effect with fewer trials.
```

### Speed vs. size/shape
#### Speed vs. GS
```{r}
# Speed vs. gosner stage, by treatment
stats %>% filter(!is.na(gs), crop == "First half second") %>% 
  ggplot(aes(x = jitter(gs), y = sline_speed_m_s, col = treatment))+
  facet_wrap(~treatment)+
  geom_point() +
  geom_smooth(method = "lm")
```

#### Speed vs. shape
```{r}
# Speed vs. PC1
stats %>% filter(crop == "First half second") %>%
  ggplot(aes(x = PC1, y = sline_speed_m_s, col = treatment))+
  facet_wrap(~treatment)+
  geom_point()+
  geom_smooth(method = "lm")
# Seems to be very little relationship between speed and PC1

# Speed vs. PC2
stats %>% filter(crop == "First half second") %>%
  ggplot(aes(x = PC2, y = sline_speed_m_s, col = treatment))+
  facet_wrap(~treatment)+
  geom_point()+
  geom_smooth(method = "lm")
# A bit more of an effect for PC2. 
```

## Join water temp data
### Load and prep data
```{r}
expdata <- read.csv("data/inputs/Burst_Speed_Data.csv")

expdata <- expdata %>% mutate(Date = as.character(Date),
                              year = 2019,
                              month = as.numeric(substr(Date, 1, 1)),
                              day = as.numeric(str_extract(Date, pattern = "(?<=^[[:digit:]]\\/)[[:digit:]]{1,2}(?=\\/)")),
                              Treatment = capwords(tolower(Treatment)))
expdata <- expdata %>% mutate(tadpole = case_when(Clutch != "W" ~ 
                                                    paste0(Pond, "_0", Clutch, "_", Indiv, "_", Treatment),
                                                  Clutch == "W" &  nchar(as.character(Indiv)) == 1 ~ 
                                                    paste0(Pond, "_W_0", Indiv, "_Wild"),
                                                  Clutch == "W" & nchar(as.character(Indiv)) == 2 ~
                                                    paste0(Pond, "_W_", Indiv, "_Wild")))
```

### Join to burst speed stats
```{r}
expdata_tojoin <- expdata %>% select(T_initial, T_final, tadpole, year, month, day, Cam) %>% mutate(Cam = factor(as.character(Cam))) %>% filter(!is.na(T_initial))

stats <- stats %>% left_join(expdata_tojoin, by = c("tadpole", "burst_year" = "year", "burst_month" = "month", "burst_day" = "day", "camera" = "Cam"))
```

### Investigate problems
```{r}
na_rows <- stats %>% filter(is.na(T_initial))
shouldbena <- expdata %>% filter(is.na(T_initial)) %>% select(tadpole) %>% unique()
problematic_na_rows <- na_rows %>% filter(!tadpole %in% shouldbena$tadpole) %>% group_by(tadpole) %>% slice(1) %>% select(tadpole, burst_year, burst_month, burst_day, camera, T_initial)

expdata_tojoin %>% filter(tadpole %in% problematic_na_rows$tadpole)

# There are a few problems left, but it's not due to a problematic join. We just seem to be missing the temp data for these. For the purposes of figuring out whether burst speed relates to water temp, we don't need to worry about that.

stats <- stats %>% mutate(T_final = NULL) # remove T_final, since we don't have that for most of them.
```

## Effects of water temp
```{r}
stats %>% ggplot(aes(x = T_initial, y = sline_speed_m_s, col = treatment))+
  geom_point()+
  geom_smooth(method = "lm")
```

## Partition data
### Only the first half second
```{r}
stats <- stats %>% filter(crop == "First half second")
```

### Fastest 1 trial
```{r}
stats_fast1 <- stats %>% group_by(tadpole) %>% filter(!is.na(sline_speed_m_s)) %>%
  filter(sline_speed_m_s == max(sline_speed_m_s))
```

### Fastest 3 trials
```{r}
stats_fast3 <- stats %>% filter(!is.na(sline_speed_m_s)) %>% group_by(tadpole) %>% arrange(desc(sline_speed_m_s), .by_group = T) %>% slice(1:3) 

stats_fast3_avg <- stats_fast3 %>% select(-c(trial, T_initial)) %>% summarize_if(.predicate = is.numeric, .funs = mean)
```

## Output
```{r}
save(stats_fast1, file = "data/outputs/stats_fast1.Rda")
save(stats_fast3, file = "data/outputs/stats_fast3.Rda")
save(stats_fast3_avg, file = "data/outputs/stats_fast3_avg.Rda") 
```

