---
title: "JEZ_Analysis_2"
author: "Kaija Gahm"
date: "4/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries
```{r}
source("libraries.R")
```

# Load data
```{r}
# Lab and wild
load("data/outputs/dat3_nonas.Rda")
load("data/outputs/wild_nonas.Rda")

# Oviposition dates
load("data/outputs/ovip_dates.Rda")
```

# Calculate developmental rates (stages/day)
## Wild
### Calculate
```{r}
wild <- wild_nonas %>%
  mutate(devrate = (gs-1)/as.numeric((burst_date-ovip_date)))
```

## Lab
### Bind oviposition dates
```{r}
lab <- ovip_dates %>%
  right_join(dat3_nonas, by = "pond")
```
### Calculate
```{r}
lab <- lab %>%
  mutate(devrate = (gs-1)/as.numeric((burst_date-ovip_date)))
```

Hypothesis: Developing faster correlates with reduced performance
We did a development manipulation by varying temperature
# Do dev rates differ by temperature treatment? (and wild)
### Prepare data
```{r}
wilddev <- wild %>% 
  select(devrate, tadpole) %>%
  mutate(treatment = "Wild")

labdev <- lab %>%
  select(devrate, tadpole, treatment)

dev <- bind_rows(wilddev, labdev)
```

## Dev boxplot (lab)
```{r}
labdev %>%
  group_by(treatment, tadpole) %>%
  summarize(devrate = mean(devrate)) %>%
  ungroup() %>%
  mutate(treatment = fct_relevel(treatment, c("Low", "High"))) %>%
  ggplot(aes(x = treatment, desc, y = devrate, fill = treatment))+
  geom_boxplot()+
  scale_fill_manual(values = c("dodgerblue3", "orange"))+
  theme_minimal()+
  theme(text = element_text(size = 22), legend.position = "none")+
  ylab("Developmental rate (stages/day)")+
  xlab("Temperature treatment")
```

### ANOVA
```{r}
mod <- lm(devrate ~ treatment, data = dev)
summary(mod)

hdev <- coef(mod)[1]
ldev <- coef(mod)[1] + coef(mod)[2]
wdev <- coef(mod)[1] + coef(mod)[3]
```
YES, they differ by treatment.
Average dev rate for high was 0.668914 gs/d, for low was 0.668914-0.267409 = 0.401505 gs/d, and for wild was 0.668914-0.167634 = 0.50128 gs/d.

## Quantify: dev rate increase per degree celsius
### Lab
```{r}
mod <- lm(devrate ~ avg_temp, data = lab)
summary(mod)
```
We observed an 0.086 gs/d increase in development rate for an incubator temperature increase of 1ºC.

### Wild
```{r}
mod <- lm(devrate ~ avg_pond_temp, data = wild)
summary(mod)
```
0.019 gs/d increase in dev rate for a pond temperature increase of 1ºC.

# Does performance vary with developmental rate?
## LAB
### Mixed model
```{r}
performance_mod_lab <- lmer(sline_speed_mm_s ~ mass_at_stage * devrate + gs + PC2 + T_initial + (1 | pond/clutch/tadpole), data = lab, REML = T)

summary(performance_mod_lab)

# Get R-squared
MuMIn::r.squaredGLMM(performance_mod_lab)

# Confidence intervals for the coefficients
#ci_lab <- confint.merMod(performance_mod_lab, method = "boot", nsim = 1000, oldNames = FALSE)
#ci_lab
```
YES, performance is negatively correlated with temperature treatment.

### Partial regression plots
#### Make data (mean-constant except for treatment and devrate)
```{r}
# make new data
newlab.devrate <- lab %>%
  select(mass_at_stage, devrate, gs, PC2, T_initial, pond, clutch, tadpole, treatment) %>%
  mutate(mass_at_stage = mean(mass_at_stage),
         gs = median(gs),
         PC2 = mean(PC2),
         T_initial = mean(T_initial))

# Add predictions
newlab.devrate <- newlab.devrate %>%
  mutate(fit.re = predict(performance_mod_lab, newdata = newlab.devrate, re.form  = NULL),
         resid = summary(performance_mod_lab)$resid,
         fit = fit.re + resid) %>%
  select(pond, clutch, tadpole, treatment, devrate, fit)
```

#### Tradeoff plot
```{r}
# Main plot
p <- newlab.devrate %>%
  group_by(treatment, pond, clutch, tadpole) %>%
  summarize(devrate = mean(devrate), fit = mean(fit)) %>%
  group_by(treatment, pond, clutch) %>%
  summarize(devrate = mean(devrate), fit = mean(fit)) %>%
  group_by(treatment, pond) %>%
  summarize(devrate = mean(devrate), fit = mean(fit)) %>%
  ggplot(aes(x = devrate, y = fit, col = treatment))+
  geom_line(aes(group = pond), color = "black", lwd = 0.8, alpha = 0.7)+
  geom_point(size = 3, alpha = 0.7)+
  scale_color_manual(name = "Temperature treatment", 
                     values = c("orange", "dodgerblue3"))+
  theme_minimal()+
  theme(legend.position = "bottom")+
  xlab("Developmental rate (GS/d)")+
  ylab("Predicted burst speed (mm/s)")
p

# Marginal plots
xdensity <- axis_canvas(p, axis = "x")+
  geom_density(data = newlab.devrate %>%
                 group_by(treatment, pond, clutch, tadpole) %>%
                 summarize(devrate = mean(devrate), fit = mean(fit)),
               aes(x = devrate, fill = treatment, col = treatment),
               alpha = 0.3)+
  scale_fill_manual(values = c("orange", "dodgerblue3"))+
  scale_color_manual(values = c("orange", "dodgerblue3"))
xdensity

combined_x <- insert_xaxis_grob(p, xdensity, position = "top")
ggdraw(combined_x)

ydensity <- axis_canvas(p, axis = "y", coord_flip = TRUE) +
  geom_density(data = newlab.devrate %>%
                 group_by(treatment, pond, clutch, tadpole) %>%
                 summarize(devrate = mean(devrate), fit = mean(fit)),
               aes(x = fit, fill = treatment, col = treatment),
               alpha = 0.3)+
  scale_fill_manual(values = c("orange", "dodgerblue3"))+
  scale_color_manual(values = c("orange", "dodgerblue3"))+
  coord_flip()

combined_y <- insert_yaxis_grob(p, ydensity, position = "right")
ggdraw(combined_y)

combined_plot <- combined_x %>%
  insert_yaxis_grob(., ydensity, position = "right")
ggdraw(combined_plot)
```

#### Facetted by treatment
THIS IS BAD BECAUSE THE SCALES ARE WRONG! Need to be able to have the limits shift but the scale stay the same. hmm...
```{r}
newlab.devrate %>%
  group_by(treatment, pond, clutch, tadpole) %>%
  summarize(devrate = mean(devrate), fit = mean(fit)) %>%
  ggplot(aes(x = devrate, y = fit, col = treatment))+
  geom_point(alpha = 0.7)+
  scale_color_manual(name = "Temperature treatment", 
                     values = c("orange", "dodgerblue3"))+
  theme_minimal()+
  geom_smooth(method = "lm")+
  facet_wrap(~treatment, scales = "free_x")+
  theme(legend.position = "bottom")+
  xlab("Developmental rate (GS/d)")+
  ylab("Predicted burst speed (mm/s)")
```

#### Both treatments
```{r}
newlab.devrate %>%
  group_by(treatment, pond, clutch, tadpole) %>%
  summarize(devrate = mean(devrate), fit = mean(fit)) %>%
  ggplot(aes(x = devrate, y = fit))+
  geom_smooth(method = "lm", se = F, col = "black", lty = "dashed")+
  geom_point(alpha = 0.7, aes(col = treatment))+
  scale_color_manual(name = "Temperature treatment", 
                     values = c("orange", "dodgerblue3"))+
  scale_fill_manual(name = "Temperature treatment", 
                     values = c("orange", "dodgerblue3"))+
  theme_minimal()+
  geom_smooth(method = "lm", aes(col = treatment, fill = treatment))+
  theme(legend.position = "bottom")+
  xlab("Developmental rate (GS/d)")+
  ylab("Predicted burst speed (mm/s)")
```
YES, tradeoff between developmental rate and performance.

## WILD
### Mixed model
```{r}
performance_mod_wild <- lmer(sline_speed_mm_s ~ mass_at_stage * devrate + gs + PC2 + T_initial + (1 | pond/tadpole), data = wild, REML = T)

summary(performance_mod_wild)

# Get R-squared
MuMIn::r.squaredGLMM(performance_mod_wild)

# Confidence intervals for the coefficients
#ci_wild <- confint.merMod(performance_mod_wild, method = "boot", nsim = 1000, oldNames = FALSE)
#ci_wild
```

### Partial regression plots
#### Make data (mean-constant except for treatment and devrate)
```{r}
# make new data
newwild.devrate <- wild %>%
  select(mass_at_stage, devrate, gs, PC2, T_initial, pond, tadpole) %>%
  mutate(mass_at_stage = mean(mass_at_stage),
         gs = median(gs),
         PC2 = mean(PC2),
         T_initial = mean(T_initial))

# Add predictions
newwild.devrate <- newwild.devrate %>%
  mutate(fit.re = predict(performance_mod_wild, newdata = newwild.devrate, re.form  = NULL),
         resid = summary(performance_mod_wild)$resid,
         fit = fit.re + resid) %>%
  select(pond, tadpole, devrate, fit)
```

#### Tradeoff plot
```{r}
w <- newwild.devrate %>%
  group_by(pond, tadpole) %>%
  summarize(devrate = mean(devrate), fit = mean(fit)) %>%
  group_by(pond) %>%
  summarize(devrate = mean(devrate), fit = mean(fit)) %>%
  ggplot(aes(x = devrate, y = fit))+
  geom_point(size = 2, alpha = 0.9)+
  theme_minimal()+
  geom_smooth(data = newwild.devrate %>%
                group_by(pond, tadpole) %>%
                summarize(devrate = mean(devrate), fit = mean(fit)),
              method = "lm", col = "black")+
  xlab("Developmental rate (GS/d)")+
  ylab("Predicted burst speed (mm/s)")
w

newwild.devrate %>%
  group_by(pond, tadpole) %>%
  summarize(devrate = mean(devrate), fit = mean(fit)) %>%
  ggplot(aes(x = devrate, y = fit))+
  geom_point(size = 3, alpha = 0.7)+
  theme_minimal()+
  geom_smooth(method = "lm", col = "black")+
  xlab("Developmental rate (stages/day)")+
  ylab("Burst speed (mm/sec)")+
  theme(text = element_text(size = 22))

# Marginal plots
xdensity <- axis_canvas(w, axis = "x")+
  geom_density(data = newwild.devrate %>%
                 group_by(pond, tadpole) %>%
                 summarize(devrate = mean(devrate), fit = mean(fit)),
               aes(x = devrate),
               alpha = 0.3, fill = "black")
xdensity

combined_x <- insert_xaxis_grob(w, xdensity, position = "top")
ggdraw(combined_x)

ydensity <- axis_canvas(w, axis = "y", coord_flip = TRUE) +
  geom_density(data = newwild.devrate %>%
                 group_by(pond, tadpole) %>%
                 summarize(devrate = mean(devrate), fit = mean(fit)),
               aes(x = fit),
               alpha = 0.3, fill = "black")+
  coord_flip()


combined_y <- insert_yaxis_grob(w, ydensity, position = "right")
ggdraw(combined_y)

combined_plot <- combined_x %>%
  insert_yaxis_grob(., ydensity, position = "right")
ggdraw(combined_plot)
```
YES, tradeoff between developmental rate and performance.

#### Facetted by pond
```{r}
# model by pond
bypond <- lm(fit ~ devrate * pond, data = newwild.devrate %>% 
               group_by(pond, tadpole) %>% 
               summarize(devrate = mean(devrate), fit = mean(fit)))
summary(bypond)
visreg(bypond)

newwild.devrate %>%
  group_by(pond, tadpole) %>%
  summarize(devrate = mean(devrate), fit = mean(fit)) %>%
  ggplot(aes(x = devrate, y = fit))+
  geom_point(size = 1.5, alpha = 0.7)+
  theme_minimal()+
  geom_smooth(method = "lm", col = "black", lwd = 0.7)+
  xlab("Developmental rate (GS/d)")+
  ylab("Predicted burst speed (mm/s)")+
  facet_wrap(~pond)+
  theme(text = element_text(size = 22))
```

# Compare wild and lab: burst speeds
```{r}
partialreg_data <- bind_rows(newwild.devrate %>% mutate(treatment = "Wild"), newlab.devrate) %>%
  group_by(treatment, tadpole) %>%
  summarize(devrate = mean(devrate),
            fit = mean(fit))

pc2_data <- dat3_nonas %>%
  select(tadpole, treatment, PC2) %>%
  distinct() %>%
  bind_rows(wild_nonas %>%
              select(tadpole, PC2) %>%
              mutate(treatment = "Wild") %>%
              distinct())

head(pc2_data)
head(partialreg_data)

partialreg_data %>%
  ggplot(aes(x = devrate, y = fit, col = treatment))+
  geom_point(alpha = 0.8)+
  geom_smooth(method = "lm")+
  theme_minimal()+
  scale_color_manual(name = "Temperature treatment", values = c("orange", "dodgerblue3", "black"))+
  theme(legend.position = "bottom")+
  ylab("Burst speed (mm/sec)")+
  xlab("Developmental rate (stages/day)")

partialreg_data %>%
  ggplot(aes(x = treatment, y = fit))+
  geom_boxplot()
```




# Allocation
Allocation rates are different between lab and wild.
```{r}
load("data/outputs/dat3_nonas.Rda")
load("data/outputs/wild_nonas.Rda")
lab_alloc <- dat3_nonas %>%
  select(gs, mass, treatment, tadpole) %>%
  distinct()

alloc_mod_lab <- lm(mass ~ gs + treatment, data = lab_alloc)
summary(alloc_mod_lab)

wild_alloc <- wild_nonas %>%
  select(gs, mass, tadpole) %>%
  mutate(treatment = "Wild") %>%
  distinct()

alloc_mod_wild <- lm(mass ~ gs, data = wild_alloc)
summary(alloc_mod_wild)

alloc <- bind_rows(lab_alloc, wild_alloc)

plot.alloc <- alloc %>% ggplot(aes(x = jitter(gs, amount = 0.15), y = mass, col = treatment))+ 
  geom_point(alpha = 0.7)+
  geom_smooth(method = "lm", aes(fill = treatment))+
  scale_color_manual(name = "Temperature treatment", values = c("orange", "dodgerblue3", "black"))+
  scale_fill_manual(name = "Temperature treatment", values = c("orange", "dodgerblue3", "black"))+
  theme_minimal()+
  theme(legend.position = "bottom")+
  ylab("Mass (g)")+
  xlab("Gosner stage")
plot.alloc
```

# Performance plasticity
## Create data
```{r}
# new model including treatment instead of dev rate
pls.model <- lmer(sline_speed_mm_s ~ mass_at_stage * treatment + gs + PC2 + T_initial + (1 | pond/clutch/tadpole), data = lab)

newlab.pls <- lab %>%
  select(mass_at_stage, treatment, gs, PC2, T_initial, 
         pond, clutch, tadpole) %>%
  mutate(mass_at_stage = mean(mass_at_stage),
         gs = median(gs),
         PC2 = mean(PC2),
         T_initial = mean(T_initial))

# Add predictions
newlab.pls <- newlab.pls %>%
  mutate(fit.re = predict(pls.model, newdata = newlab.pls, re.form  = NULL),
         resid = summary(pls.model)$resid,
         fit = fit.re + resid) %>%
  select(treatment, pond, clutch, tadpole, fit)
```

## Individual means
```{r}
indiv_means <- newlab.pls %>%
  group_by(treatment, pond, clutch, tadpole) %>%
  summarize(indiv_mean = mean(fit))
```

## Clutch means
```{r}
clutch_means <- indiv_means %>%
  group_by(treatment, pond, clutch) %>%
  summarize(clutch_mean = mean(indiv_mean))

clutch_means %$% table(clutch)
```

#### Remove clutches not in both treatments
```{r}
clutch_means <- clutch_means %>% 
  group_by(clutch) %>%
  mutate(freq = n()) %>%
  ungroup() %>%
  filter(freq == 2) %>%
  select(-freq)
```

### Treatment differences
```{r}
treatment_diffs <- clutch_means %>%
  group_by(clutch) %>%
  summarize(diff = clutch_mean[treatment == "Low"] - 
              clutch_mean[treatment == "High"])

treatment_diffs %>% ggplot(aes(x = diff))+
  geom_histogram()
```
What I've calculated here is D, the "scope of plastic response" (Stearns 1992, cited in Valladares _et al._ 2006): mean at low resource availability - mean at high resource availability. (note low vs. high is switched from the paper because in this case high temp ~ slower speed.)

## Individuals vs. means
### Opposite treatment means
```{r}
clutch_means <- clutch_means %>%
  group_by(clutch) %>%
  mutate(opposite_clutch_mean = case_when(treatment == "High" ~ 
                                            clutch_mean[treatment == "Low"],
                                   treatment == "Low" ~ 
                                     clutch_mean[treatment == "High"]))
```

### Join to indivs
```{r}
indiv_means <- indiv_means %>%
  left_join(clutch_means, by = c("treatment", "clutch", "pond"))

head(indiv_means)
```

### Calculate D
If we always subtract individual - mean, or always mean - individual, the distribution will be centered around 0 because we'll be subtracting a mixture of high-low and low-high. Need to be consistent with temperatures, which means inconsistent with direction of subtraction.

We're going to default to always subtracting low - high. 
```{r}
indiv_means <- indiv_means %>%
  mutate(indiv_opp_clutch_diff = 
           case_when(treatment == "High" ~ 
                       opposite_clutch_mean - indiv_mean,
                     treatment == "Low" ~ 
                       indiv_mean - opposite_clutch_mean))
```

#### Clean data
```{r}
plasticity <- indiv_means %>% select(-c(clutch_mean)) %>% rename(D = indiv_opp_clutch_diff)

# add a column for just the clutch
plasticity <- plasticity %>% mutate(clutch_only = factor(str_extract(clutch, pattern = "(?<= )[[:digit:]](?=$)")))
save(plasticity, file = "data/outputs/plasticity.Rda")
```

## Plot
```{r}
# Get order of ponds by pond temperature
load("data/outputs/pond_avg_temps.Rda")

# Get order of ponds by avg temp
temp_pond_order <- pond_avg_temps %>%
  arrange(-avg_pond_temp) %>%
  pull(pond) %>%
  as.character()

# Make plot
## Define helper functions
reorder_within <- function(x, by, within, fun = mean, sep = "___", ...) {
  new_x <- paste(x, within, sep = sep)
  stats::reorder(new_x, by, FUN = fun)
}

scale_x_reordered <- function(..., sep = "___") {
  reg <- paste0(sep, ".+$")
  ggplot2::scale_x_discrete(labels = function(x) gsub(reg, "", x), ...)
}

## make the plot
plasticity %>%
  ungroup() %>%
  mutate(pond = fct_relevel(pond, temp_pond_order)) %>%
  filter(!is.na(clutch_only)) %>%
  ggplot(aes(x = reorder_within(clutch_only, -D, pond, mean), y = D))+
  geom_rect(aes(fill = pond),
            xmin = -Inf, xmax = Inf,
            ymin = -Inf, ymax = Inf, 
            alpha = 0.3)+
  geom_hline(aes(yintercept = 0), color = "black")+
  scale_fill_manual(values = rep(c("grey99", "grey78"), 5))+
  geom_boxplot()+
  scale_x_reordered()+
  facet_grid(cols = vars(pond), scales = "free_x")+
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    legend.position = "none"
  )+
  ylab("Plasticity in performance: low - high temp (mm/s)")+
  xlab("Pond")
```

## Differences in plasticity by pond?
```{r}
# Linear model, assuming roughly equal sample sizes per pond, which we have
table(plasticity$pond) # roughly equal sample sizes--this should be fine.
pls.mod <- lm(D ~ pond, data = plasticity)
summary(pls.mod) # A few are significantly different w/r/t plasticity: BO, GB, and (marginally) PB.
```

# Dev rate plasticity
## Data
```{r}
head(lab)
```

## Individual means
```{r}
indiv_means_dev <- lab %>%
  group_by(treatment, pond, clutch, tadpole) %>%
  summarize(indiv_mean_dev = mean(devrate))
```

## Clutch means
```{r}
clutch_means_dev <- indiv_means_dev %>%
  group_by(treatment, pond, clutch) %>%
  summarize(clutch_mean_dev = mean(indiv_mean_dev))

clutch_means_dev %$% table(clutch)
```

#### Remove clutches not in both treatments
```{r}
clutch_means_dev <- clutch_means_dev %>% 
  group_by(clutch) %>%
  mutate(freq = n()) %>%
  ungroup() %>%
  filter(freq == 2) %>%
  select(-freq)
```

### Treatment differences
```{r}
treatment_diffs_dev <- clutch_means_dev %>%
  group_by(clutch) %>%
  summarize(diff = clutch_mean_dev[treatment == "Low"] - 
              clutch_mean_dev[treatment == "High"])

treatment_diffs_dev %>% ggplot(aes(x = diff))+
  geom_histogram()
```
This is plasticity in _developmental rates_, GS/day.

## Individuals vs. means
### Opposite treatment means
```{r}
clutch_means_dev <- clutch_means_dev %>%
  group_by(clutch) %>%
  mutate(opposite_clutch_mean_dev = case_when(treatment == "High" ~ 
                                            clutch_mean_dev[treatment == "Low"],
                                   treatment == "Low" ~ 
                                     clutch_mean_dev[treatment == "High"]))
```

### Join to indivs
```{r}
indiv_means_dev <- indiv_means_dev %>%
  left_join(clutch_means_dev, by = c("treatment", "clutch", "pond"))

head(indiv_means_dev)
```

### Calculate D
If we always subtract individual - mean, or always mean - individual, the distribution will be centered around 0 because we'll be subtracting a mixture of high-low and low-high. Need to be consistent with temperatures, which means inconsistent with direction of subtraction.

We're going to default to always subtracting low - high. 
```{r}
indiv_means_dev <- indiv_means_dev %>%
  mutate(indiv_opp_clutch_diff_dev = 
           case_when(treatment == "High" ~ 
                       opposite_clutch_mean_dev - indiv_mean_dev,
                     treatment == "Low" ~ 
                       indiv_mean_dev - opposite_clutch_mean_dev))
```

#### Clean data
```{r}
pls_dev <- indiv_means_dev %>% select(-c(clutch_mean_dev)) %>% rename(D = indiv_opp_clutch_diff_dev)

# add a column for just the clutch
pls_dev <- pls_dev %>% mutate(clutch_only = factor(str_extract(clutch, pattern = "(?<= )[[:digit:]](?=$)")))
save(pls_dev, file = "data/outputs/pls_dev.Rda")
```

## Plot
```{r}
## make the plot
pls_dev %>%
  ungroup() %>%
  mutate(pond = fct_relevel(pond, temp_pond_order)) %>%
  filter(!is.na(clutch_only)) %>%
  ggplot(aes(x = reorder_within(clutch_only, -D, pond, mean), y = D))+
  geom_rect(aes(fill = pond),
            xmin = -Inf, xmax = Inf,
            ymin = -Inf, ymax = Inf, 
            alpha = 0.3)+
  # geom_hline(aes(yintercept = 0), color = "black")+
  scale_fill_manual(values = rep(c("grey99", "grey78"), 5))+
  geom_boxplot()+
  scale_x_reordered()+
  facet_grid(cols = vars(pond), scales = "free_x")+
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    legend.position = "none"
  )+
  ylab("Plasticity in development rate: low - high temp (GS/d)")+
  xlab("Pond")
```

# Dev rate plasticity vs. performance plasticity
```{r}
head(plasticity)
head(pls_dev)

plss <- plasticity %>% 
  select(treatment, pond, clutch, D) %>%
  rename(D_perf = D) %>%
  left_join(pls_dev %>% 
              select(treatment, pond, clutch, D)) %>%
  mutate(D_dev = D) %>%
  select(-D)

head(plss)

plss %>%
  # group_by(pond, clutch) %>%
  # summarize(D_dev = mean(D_dev),
  #           D_perf = mean(D_perf)) %>%
  ggplot(aes(x = D_dev, y = D_perf))+
  geom_point(aes(col = pond))+
  theme_minimal()+
  xlab("Developmental plasticity")+
  ylab("Performance plasticity")
```


# Descriptive stats
## Make data
```{r}
lab_covars <- lab %>% 
  select(gs, PC2, T_initial, treatment)

wild_covars <- wild %>%
  select(gs, PC2, T_initial) %>%
  mutate(treatment = "Wild")

covars <- bind_rows(lab_covars, wild_covars)

head(covars)
```

## Boxplots
```{r}
# Gosner stage
gs.plot <- covars %>%
  ggplot(aes(x = treatment, y = gs, fill = treatment))+
  geom_boxplot()+
  theme_minimal()+
  scale_fill_manual(values = c("orange", "dodgerblue3", "gray60"))+
  theme(legend.position = "none", axis.title.x = element_blank())+
  ylab("Gosner stage")
gs.plot

# PC2
pc2.plot <- covars %>%
  ggplot(aes(x = treatment, y = PC2, fill = treatment))+
  geom_boxplot()+
  theme_minimal()+
  scale_fill_manual(values = c("orange", "dodgerblue3", "gray60"))+
  theme(legend.position = "none", axis.title.x = element_blank())+
  ylab("PC2")
pc2.plot

# T_initial
t.plot <- covars %>%
  ggplot(aes(x = treatment, y = T_initial, fill = treatment))+
  geom_boxplot()+
  theme_minimal()+
  scale_fill_manual(values = c("orange", "dodgerblue3", "gray60"))+
  theme(legend.position = "none", axis.title.x = element_blank())+
  ylab("Tray temperature")
t.plot

plot_grid(gs.plot, pc2.plot, t.plot)

```

# Morpho plot
```{r}
load("data/outputs/morpho_scores.Rda")
head(morpho_scores)

morpho_scores %>%
  ggplot(aes(x = PC1, y = PC2, col = treatment))+
  geom_point(alpha = 0.8)+
  scale_color_manual(name = "Temperature treatment", values = c("orange", "dodgerblue3", "black"))+
  theme_minimal()+
  theme(legend.position = "bottom")+
  ylim(c(-0.15, 0.15))+
  xlim(c(-0.15, 0.25))

morpho_scores %>%
  ggplot(aes(x = PC2, col = treatment, fill = treatment))+
  geom_density(alpha = 0.6)+
  scale_fill_manual(name = "Temperature treatment", values = c("orange", "dodgerblue3", "black")) +
  scale_color_manual(name = "Temperature treatment", values = c("orange", "dodgerblue3", "black"))+
  theme_minimal()+
  theme(legend.position = "bottom", text = element_text(size = 22))+
  ylab("")
  
```

# Performance vs. pond temperature (wild)
## Load temperatures
```{r}
load("data/outputs/pond_avg_temps.Rda")
head(newwild.devrate)

wild_temps <- left_join(newwild.devrate, pond_avg_temps, by = "pond")

wild_temps

wild_temps %>%
  ggplot(aes(x = avg_pond_temp))
```

## Plot
```{r}
wild_temps %>%
  group_by(avg_pond_temp, tadpole) %>%
  summarize(fit = mean(fit)) %>%
  ggplot(aes(x = jitter(avg_pond_temp, amount = 0.03), y = fit))+
  geom_point(alpha = 0.7)+
  geom_smooth(method = "lm", col = "black")+
  theme_minimal()+
  ylab("Predicted burst speed (mm/s)")+
  xlab("2019 mean daily pond temperature (ºC)")
```

# Dev rate vs. pond temperature
## Plot
```{r}
wild_temps %>%
  group_by(avg_pond_temp, tadpole) %>%
  summarize(devrate = mean(devrate)) %>%
  ggplot(aes(x = jitter(avg_pond_temp, amount = 0.03), y = devrate))+
  geom_point(alpha = 0.7)+
  geom_smooth(method = "lm", col = "black")+
  theme_minimal()+
  ylab("Developmental rate (GS/d)")+
  xlab("2019 mean daily pond temperature (ºC)")
```


