---
title: "dev_growth_rates"
author: "Kaija Gahm"
date: "2/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load libraries
```{r, include = F}
source("libraries.R")
source("process_coordinates_funs.R")
```

# Read data
```{r}
dev <- read.csv("data/inputs/dev_measurements_kaija_andis.csv", as.is = T)
```

# Data prep
## Clean data
```{r}
load("data/outputs/dat3.Rda")
head(dat3)

# Make tadpole column
dev <- dev %>% mutate(tadpole = paste0(pond, "_0", clutch, "_", indiv, "_", treatment))

# Read dates as dates
dev <- dev %>% mutate(photo.date = mdy(photo.date))

dev <- dev %>% mutate(person = paste0(person, "_orig"),
                      person_verify = paste0(person_verify, "_verify")) 

# clean ponds
table(dev$pond)
dev$pond[dev$pond == "DB"] <- "PB" # assuming it's PB, not DE?
```

## Check between-person agreement
```{r}
# Check error
checks <- dev %>% filter(!is.na(stage.verify)) # get only the ones that were staged a second time
checks <- checks %>% mutate(stagediff = stage - stage.verify)

checks %>% ggplot(aes(x = stagediff))+
  geom_density(fill = "darkgrey")+ 
  ggtitle("Difference in stages")+
  facet_grid(rows = vars(person), cols = vars(person_verify))

# SVL
checks %>% ggplot(aes(x = svl.mm - svl.mm.verify))+
  geom_density(fill = "darkgrey")+ # oof 
  ggtitle("Difference in SVL")+
  facet_grid(rows = vars(person), cols = vars(person_verify))

checks %>% ggplot(aes(x = svl.mm - svl.mm.verify))+
  geom_density(aes(fill = person_verify, col = person_verify), alpha = 0.5)+
  ggtitle("Difference in SVL")+
  facet_wrap(~person)

# Total length
checks %>% ggplot(aes(x = totl.mm - totl.mm.verify))+
  geom_density(fill = "darkgrey")+ # oof 
  ggtitle("Difference in total length")+
  facet_grid(rows = vars(person), cols = vars(person_verify))

checks %>% ggplot(aes(x = totl.mm - totl.mm.verify))+
  geom_density(aes(fill = person_verify, col = person_verify), alpha = 0.5)+
  ggtitle("Difference in total length")+
  facet_wrap(~person)
```

## Align Andis/Kaija
```{r}
# Subtract 0.25 from Kaija's SVL measurements
dev <- dev %>%
  mutate(svl.mm = case_when(person == "kaija_orig" ~ svl.mm - 0.25,
                            TRUE ~ svl.mm),
         aligned = case_when(person == "kaija_orig" ~ TRUE,
                             TRUE ~ FALSE))

# Do the same to Checks
checks <- checks %>%
  mutate(svl.mm = case_when(person == "kaija_orig" ~ svl.mm - 0.25,
                            TRUE ~ svl.mm),
         aligned = case_when(person == "kaija_orig" ~ TRUE,
                             TRUE ~ FALSE))

# re-do the above plot. 
checks %>% ggplot(aes(x = svl.mm - svl.mm.verify))+
  geom_density(aes(fill = person_verify, col = person_verify), alpha = 0.5)+
  ggtitle("Difference in SVL")+
  facet_wrap(~person)
```

# Visualize
## Rough dev graphs
```{r}
head(dev)
dev %>% filter(!is.na(pond), !is.na(stage)) %>%
  ggplot(aes(x = photo.date, y = stage, col = treatment, group = tadpole))+
  geom_point(size = 0.5)+
  geom_line(lwd = 0.5, alpha = 0.1)+
  theme_minimal()+
  guides(color = guide_legend(override.aes = list(size=5)))+
  xlab("Date")+
  ylab("Gosner stage")

dev %>% filter(!is.na(pond), !is.na(svl.mm)) %>%
  ggplot(aes(x = photo.date, y = svl.mm, col = treatment, group = tadpole))+
  geom_point(size = 0.5)+
  geom_line(lwd = 0.5, alpha = 0.1)+
  theme_minimal()+
  guides(color = guide_legend(override.aes = list(size=5)))+
  xlab("Date")+
  ylab("SVL (mm)")

dev %>% filter(!is.na(pond), !is.na(totl.mm)) %>%
  ggplot(aes(x = photo.date, y = totl.mm, col = treatment, group = tadpole))+
  geom_point(size = 0.5)+
  geom_line(lwd = 0.5, alpha = 0.1)+
  theme_minimal()+
  guides(color = guide_legend(override.aes = list(size=5)))+
  xlab("Date")+
  ylab("Total length (mm)")

# Summarize means by pond
pondmeans <- dev %>% 
  group_by(treatment, pond, photo.date) %>% 
  summarize(mstage = mean(stage, na.rm = T),
            sd.stage = sd(stage, na.rm = T),
            msvl = mean(svl.mm, na.rm = T),
            sd.svl = sd(svl.mm, na.rm = T),
            mtotl = mean(totl.mm, na.rm = T),
            sd.totl = sd(totl.mm, na.rm = T))
head(pondmeans)

pondmeans %>% filter(!is.na(pond)) %>% 
  ggplot(aes(x = photo.date, y = mstage, col = pond, group = interaction(treatment, pond)))+
  geom_point()+
  geom_line()+
  geom_errorbar(aes(ymin = mstage - sd.stage, ymax = mstage + sd.stage))+
  theme_minimal()

pondmeans %>% filter(!is.na(pond)) %>% 
  ggplot(aes(x = photo.date, y = msvl, col = pond, group = interaction(treatment, pond)))+
  geom_point()+
  geom_line()+
  geom_errorbar(aes(ymin = msvl - sd.svl, ymax = msvl + sd.svl))+
  theme_minimal()

pondmeans %>% filter(!is.na(pond)) %>% 
  ggplot(aes(x = photo.date, y = mtotl, col = pond, group = interaction(treatment, pond)))+
  geom_point()+
  geom_line()+
  geom_errorbar(aes(ymin = mtotl - sd.totl, ymax = mtotl + sd.totl))+
  theme_minimal()
```

# Add embryo stages
```{r}
emb <- read.csv("data/inputs/2019_Metabolic_HatchDates.csv", as.is = TRUE)

# clean hatch data
emb <- emb %>% mutate(coll.date = dmy(Coll.Date),
                          treatment = capwords(tolower(Treat))) %>%
  rename(pond = Pond,
         clutch = Clutch)

# subset for joining
emb_tojoin <- emb %>% select(pond, clutch, treatment, coll.date) %>%
  mutate(photo.date = coll.date - 1,
         stage = 1) %>%
  select(-coll.date)

# prepare for join to dev
tads <- dev %>% # create new rows for each tadpole
  select(treatment, pond, clutch, indiv, tadpole) %>% 
  distinct() %>%
  left_join(emb_tojoin, by = c("pond", "clutch", "treatment")) %>% distinct()

# join tads to dev
dev <- bind_rows(dev, tads)
tail(dev) # examine the new data we added
```

## Plot with embryo stages added
```{r}
dev %>% filter(!is.na(pond), !is.na(stage)) %>%
  ggplot(aes(x = photo.date, y = stage, col = treatment, group = tadpole))+
  geom_point(size = 0.5)+
  geom_line(lwd = 0.5, alpha = 0.1)+
  theme_minimal()+
  guides(color = guide_legend(override.aes = list(size=5)))+
  xlab("Date")+
  ylab("Gosner stage")
```

# Add final stages
```{r}
load("data/outputs/stats_fast3.Rda")
head(stats_fast3)
final_tojoin <- stats_fast3 %>% 
  select(gs, tadpole, photo_date, pond, clutch, indiv, treatment) %>% 
  mutate(photo.date = ymd(photo_date)) %>% 
  rename(stage = gs) %>%
  filter(treatment %in% c("High", "Low"))

final_tojoin %>% filter(!(tadpole %in% dev$tadpole)) # for some reason we don't have developmental data for this one tadpole, although we do have a photo at the end. What??
dev %>% filter(pond == "CPS", clutch == 6, treatment == "High")

# Join to dev
dev2 <- bind_rows(dev, final_tojoin)

tail(dev2)
```

## Plot with final stages added
```{r}
dev2 %>% filter(!is.na(pond), !is.na(stage)) %>%
  ggplot(aes(x = photo.date, y = stage, col = treatment, group = tadpole))+
  geom_point(size = 0.5)+
  geom_line(lwd = 0.5, alpha = 0.1)+
  theme_minimal()+
  guides(color = guide_legend(override.aes = list(size=5)))+
  xlab("Date")+
  ylab("Gosner stage")
```

# Add wild ones
## Get oviposition dates by pond
```{r}
# Make table of oviposition dates for each pond
# Verify that all larvae from each pond were collected on the same day
with(emb, table(pond, Coll.Date))

# Load wild data
wild_info <- read.csv("data/inputs/wild_info.csv", as.is = TRUE)

# convert from doy to date
dates <- as.Date("2019-01-01") + 0:365 # create sequence of dates
wild_info <- wild_info %>% mutate(resp_date = dates[wild_info$Resp.DOY])

# make the df
ovip_dates <- emb %>% 
  select(pond, coll.date) %>% 
  distinct() %>%
  mutate(ovip_date = coll.date - 1) %>%
  select(-coll.date)
ovip_dates

save(ovip_dates, file = "data/outputs/ovip_dates.Rda")
```

## Join wild tadpole information
```{r}
wild <- wild_info %>% select(Treat, Pond, Indiv, ID, Resp.DOY, GS)
wild <- wild %>% rename(treatment = Treat,
                        pond = Pond,
                        indiv = Indiv,
                        tadpole = ID,
                        resp.doy = Resp.DOY,
                        stage = GS) %>%
  mutate(treatment = capwords(tolower(treatment)),
         photo.date = dates[resp.doy]) %>% # call this the "photo date" in preparation for join
  select(-resp.doy) %>%
  mutate(indiv = as.character(indiv))

wild <- wild %>% left_join(ovip_dates, by = "pond") # join oviposition dates

wild_final_tojoin <- wild %>% select(treatment, pond, indiv, tadpole, photo.date, stage)

dev3 <- bind_rows(dev2, wild_final_tojoin)

wild_initial_tojoin <- wild %>% 
  select(treatment, pond, indiv, tadpole, ovip_date, stage) %>%
  rename(photo.date = ovip_date) %>% # call the oviposition date the "photo date" for purposes of joining
  mutate(stage = 1) # set stage to 1 on the oviposition date

dev4 <- bind_rows(dev3, wild_initial_tojoin)

dev4 <- dev4 %>% mutate(tadpole = case_when(treatment == "Wild" ~ paste0(tadpole, "_Wild"),
                                            TRUE ~ tadpole))
```

## Save data
```{r}
save(dev4, file = "data/outputs/dev4.Rda")
```


## Plot with wild tadpoles added
```{r}
dev4 %>% filter(!is.na(pond), !is.na(stage)) %>%
  ggplot(aes(x = photo.date, y = stage, col = treatment, group = tadpole))+
  geom_point(size = 0.5)+
  geom_line(lwd = 0.5, alpha = 0.1)+
  theme_minimal()+
  guides(color = guide_legend(override.aes = list(size=5)))+
  xlab("Date")+
  ylab("Gosner stage")

dev4 %>% filter(!is.na(pond), !is.na(stage)) %>%
  ggplot(aes(x = photo.date, y = stage, col = treatment, group = tadpole))+
  #geom_point(size = 0.5)+
  geom_smooth(se = F, method = "lm", lwd = 0.05)+
  theme_minimal()+
  guides(color = guide_legend(override.aes = list(size=5)))+
  xlab("Date")+
  ylab("Gosner stage")
```

# Pond temperatures
## Load and clean temp data
```{r}
ponds <- list.files("data/inputs/temp_logger_data_2019/", full.names = T)
templist <- lapply(ponds, read.csv, skip = 2, header = F, as.is = T)

# Set pond names
pondnames <- c("BS", "BO", "CPS", "DT", "DE", "E1", "GB", "LO", "MI", "PB")
names(templist) <- pondnames

# select columns we want
templist <- lapply(templist, select, V2, V3) # I can't believe that worked!

# remove NA's
templist <- lapply(templist, filter, !is.na(V2) & !is.na(V3))

# make names vector
namesvec <- c("datetime", "temp_c")

# set names
templist <- lapply(templist, setNames, nm = namesvec)

# parse dates and times
templist <- map(templist, mutate, datetime = parse_date_time(datetime, orders = c("mdy HM", "mdy HMS")))
lapply(templist, FUN = locate.nas)
```

## Bind together
```{r}
temps <- rbindlist(templist, use.names = T, idcol = TRUE) %>%
  rename(pond = .id)
head(temps)
```

## Crop
### Prepare logger deploy data
```{r}
deploy <- read.csv("data/inputs/logger_deploy_data.csv", as.is = T, na.strings = c("", " "))
head(deploy)

deploy <- deploy %>% 
  mutate(datetime_in = paste(date_in, time_in),
         datetime_out = paste(date_out, na.omit(time_out))) %>%
  select(datetime_in, datetime_out, pond_abbr) %>%
  filter(!is.na(pond_abbr)) %>%
  mutate(datetime_in = mdy_hm(datetime_in),
         datetime_out = mdy_hm(datetime_out))
```


### Join logger deploy data
```{r}
temps <- temps %>% left_join(deploy, by = c("pond" = "pond_abbr"))
```

### Crop by deploy data
```{r}
temps <- temps %>% 
  group_by(pond) %>%
  filter(datetime >= datetime_in & datetime < datetime_out)

## Fix deploy times for PB and GB: the recorded deploy time doesn't match what's clear on the graph. Bump it forward slightly to account for temps when the loggers were obviously still in the car.
temps <- temps %>% filter(!(pond == "PB" & datetime < "2019-03-31 16:00:00")) # fix PB deploy time
temps <- temps %>% filter(!(pond == "GB" & datetime < "2019-03-25 10:30:00")) # fix GB deploy time
```

## Add aggregators
```{r}
temps <- temps %>%
  mutate(date = substr(datetime, 1, 10),
         date_hour = substr(datetime, 1, 13)) %>%
  mutate(date = ymd(date),
         date_hour = ymd_h(date_hour))

temps <- temps %>%
  group_by(date, pond) %>% # group by day
  mutate(mean_daily_temp_c = mean(temp_c, na.rm = T)) %>% # mean daily temp
  ungroup() %>%
  group_by(date_hour, pond) %>% # group by hour
  mutate(mean_hourly_temp_c = mean(temp_c, na.rm = T)) # mean hourly temp
```

## Visualize
```{r}
temps %>% ggplot(aes(x = date, y = mean_daily_temp_c, col = pond))+
  geom_line(lwd = 1.3)+
  scale_color_viridis_d()+
  theme_minimal()+
  ylab("Temperature (ºC)")+
  xlab("Date")
```

# Get temps over the dev period
```{r}
# Oviposition and embryo collection dates for each pond
ovip_dates <- emb %>%
  select(pond, Coll.Date) %>%
  distinct() %>%
  mutate(Coll.Date = dmy(Coll.Date),
         ovip_date = Coll.Date - 1) %>%
  rename(eggmass_coll_date = Coll.Date)
head(ovip_dates)

# Get mean oviposition date for wild tadpoles
mean_ovip_date <- emb %>% 
  select(pond, Coll.Date) %>% 
  distinct() %>%
  mutate(Coll.Date = dmy(Coll.Date),
         ovip_date = Coll.Date - 1) %>%
  summarize(mean = mean(ovip_date)) %>%
  pull(mean)
mean_ovip_date

# Info about when the wild tads were collected
wild_tads <- read.csv("data/inputs/wild_info.csv", as.is = T)
wild_tads <- wild_tads %>% select(ID, Coll.DOY, Pond)

wild_tads <- wild_tads %>% 
  mutate(coll.date = as.Date(Coll.DOY, origin = "2019-01-01")) %>%
  rename(tadpole = ID,
         pond = Pond)

# Get mean collection date for wild tadpoles
mean_tad_collection_date <- wild_tads %>%
  summarize(mean = mean(coll.date)) %>%
  pull(mean)
mean_tad_collection_date
```

## Add temps for CPS
```{r}
# Get imputed CPS temps
cps_imputed <- read.csv("data/inputs/PTemps.imputed.01to19.csv", as.is = T) %>%
  rename(pond = Pond) %>%
  filter(pond == "CP") %>%
  mutate(pond = "CPS")

cps_2019 <- cps_imputed %>%
  filter(Year == "2019") %>%
  mutate(date = as.Date(DOY, origin = "2019-01-01")) %>%
  filter(date > "2019-04-02" & date < "2019-04-30") %>%
  select(pond, date, PTemp.all) %>%
  rename(mean_daily_temp_c = PTemp.all)

temps <- bind_rows(cps_2019, temps)
```

```{r}
# Calculate pond average temperatures over the mean interval (oviposition to tadpole collection)
pond_avg_temps <- temps %>%
  group_by(pond) %>%
  select(pond, date, mean_daily_temp_c) %>%
  distinct() %>%
  filter(date >= mean_ovip_date & date <= mean_tad_collection_date) %>%
  summarize(avg_pond_temp = mean(mean_daily_temp_c, na.rm = T)) %>%
  ungroup()

head(pond_avg_temps)

# Summary of all dates for wild tadpoles
wild_dates <- wild_tads %>%
  select(tadpole, coll.date, pond) %>%
  mutate(tadpole = paste0(tadpole, "_Wild")) %>%
  left_join(ovip_dates, by = "pond") %>%
  rename(tadpole_coll_date = coll.date)

save(wild_dates, file = "data/outputs/wild_dates.Rda")
save(pond_avg_temps, file = "data/outputs/pond_avg_temps.Rda")
```

# Boxplots of daily temperatures during the dev. period
```{r}
# Crop to the dev period
temps %>%
  group_by(pond) %>%
  filter(date >= mean_ovip_date & date <= mean_tad_collection_date) %>%
  ungroup() %>%
  ggplot(aes(date, mean_daily_temp_c))+
  geom_line()+
  facet_wrap(~pond)

# Make boxplots
temps %>%
  group_by(pond) %>%
  filter(!is.na(mean_daily_temp_c), date >= mean_ovip_date & date <= mean_tad_collection_date) %>%
  ungroup() %>%
  ggplot(aes(x = fct_reorder(pond, -mean_daily_temp_c), 
             y = mean_daily_temp_c))+
  geom_boxplot()+
  xlab("Pond")+
  ylab("Average daily temperature")
```

# Incubator temperatures
## Prepare data
```{r}
# Load data
incs <- read.csv("data/inputs/incubator_temps.csv", as.is = T) 

# Convert to date format
incs <- incs %>% 
  mutate(date = dmy(paste0(Date, "-2019"))) %>%
  select(-c(Date, X, X.1))

# Pivot to long
incs_long <- incs %>% pivot_longer(cols = c(high, low, mean), names_to = "pond", values_to = "temp_c")
save(incs_long, file = "data/outputs/incs_long.Rda")

# Get average incubator temperatures
## Get mean burst assay dates for each incubator. We'll use the mean oviposition date as the starting point.
mean_ovip <- "2019-04-04"

# get the mean burst dates for each incubator
load("data/outputs/dat3_nonas.Rda")
ends <- dat3_nonas %>% 
  group_by(treatment) %>%
  summarize(end = mean(burst_date))
(high_end <- ends$end[1])
(low_end <- ends$end[2])

# get high inc avg
high_avg <- incs_long %>% 
  filter(pond == "high") %>%
  filter(date >= mean_ovip & date <= high_end) %>%
  pull(temp_c) %>%
  mean(., na.rm = T)

# get low inc avg
low_avg <- incs_long %>%
  filter(pond == "low") %>%
  filter(date >= mean_ovip & date <= low_end) %>%
  pull(temp_c) %>%
  mean(., na.rm = T)

# make a df
avg_inc_temps <- data.frame(avg_temp = c(high_avg, low_avg),
                            treatment = c("High", "Low"))

# save the avg temp data
save(avg_inc_temps, file = "data/outputs/avg_inc_temps.Rda")
```

## Visualize
```{r}
incs_long %>% ggplot(aes(x = date, y = temp_c, col = pond))+
  geom_line(lwd = 1.5)
```

# Incubator and temp logger together
## Save temp data
```{r}
save(temps, file = "data/outputs/temps.Rda")
```

## Visualize
```{r}
load("data/outputs/temps.Rda")
load("data/outputs/incs_long.Rda")
load("data/outputs/wild_nonas.Rda")
load("data/outputs/dat3_nonas.Rda")

maxburst <- max(c(wild_nonas$burst_date, dat3_nonas$burst_date))

temps %>% 
  filter(!is.na(mean_daily_temp_c), 
         date >= mean_ovip_date & date <= maxburst) %>%
  ggplot(aes(x = date, y = mean_daily_temp_c, group = pond))+
  geom_line(alpha = 0.4)+
  geom_line(data = incs_long %>% filter(pond %in% c("high", "low")), 
            aes(x = date, y = temp_c, col = pond), lwd = 1.5)+
  scale_color_manual(values = c("red", "blue"))+
  xlab("Date")+
  ylab("Temperature (ºC)")

temps %>% filter(pond == "CPS")
```

## Crop to developmental intervals
```{r}
head(ovip_dates)
min_ovip <- ovip_dates %$% 
  min(ovip_date)

load("data/outputs/wild_nonas.Rda")

max_wild_coll <- wild_nonas %$%
  max(tadpole_coll_date)

min_incubator <- "2019-04-04" # This was the earliest date on which incubator temperatures were measured. Check with Andis if we need to use stocking dates instead?

# Get high and low latest burst dates
max_high_burst <- dat3_nonas %>%
  filter(treatment == "High") %$%
  max(burst_date)

max_low_burst <- dat3_nonas %>%
  filter(treatment == "Low") %$%
  max(burst_date)

# Crop incubator temps to the developmental period
incs_long_cropped <- incs_long %>%
  mutate(pond = case_when(pond == "high" ~ "High",
                               pond == "low" ~ "Low",
                               pond == "mean" ~ "Mean")) %>%
  mutate(late_date = case_when(pond == "High" ~ max_high_burst,
                               pond == "Low" ~ max_low_burst)) %>%
  filter(pond %in% c("High", "Low")) %>%
  filter(date >= min_incubator & date <= late_date)

save(incs_long_cropped, file = "data/outputs/incs_long_cropped.Rda")
```

## Visualize cropped
```{r}
load("data/outputs/loggers_cropped.Rda")
load("data/outputs/incs_long_cropped.Rda")
loggers_cropped %>% ggplot(aes(x = date, y = mean_daily_temp_c, group = pond))+
  geom_line(alpha = 0.4)+
  geom_line(data = incs_long_cropped,
            aes(x = date, y = temp_c, col = pond), lwd = 1.5)+
  scale_color_manual(values = c("red", "blue"))+
  xlab("Date")+
  ylab("Temperature (ºC)")
```


# Model dev rates
```{r}
mean_center <- function(vector){
  mu <- mean(vector, na.rm = T)
  centered_vector <- (vector-mu)
  return(centered_vector)
}

formod <- dev4 %>% 
  filter(!is.na(photo.date)) %>%
  select(treatment, photo.date, tadpole, stage, pond) %>%
  mutate(treatment = factor(treatment),
         tadpole = factor(tadpole),
         pond = factor(pond),
         photo.date = mean_center(photo.date)) %>%
  filter(treatment %in% c("High", "Low"))

dev_mod <- lmer(stage ~ photo.date * treatment + (1 | tadpole), data = formod)
summary(dev_mod)
#stage predicted by photo.date * treatment, with a random effect for tadpole. Give a brief description of this and the linear method in the methods.
```

## ...and growth rates, by SVL (mm)
```{r}
formod_growth <- dev4 %>% 
  filter(!is.na(photo.date)) %>%
  select(treatment, photo.date, tadpole, svl.mm) %>%
  mutate(treatment = factor(treatment),
         tadpole = factor(tadpole),
         photo.date = mean_center(photo.date)) %>%
  filter(treatment %in% c("High", "Low"))

growth_mod <- lmer(svl.mm ~ photo.date * treatment + (1 | tadpole), data = formod_growth)

summary(growth_mod)
```

# Dev rates by pond
```{r}
dev_mod_pond <- lmer(stage ~ as.numeric(photo.date) * treatment * pond + (1 | tadpole), data = formod)
summary(dev_mod_pond)

datesvec <- seq(min(as.numeric(formod$photo.date)), max(as.numeric(formod$photo.date)), by = 3)
testdata <- expand.grid(photo.date = datesvec,
           pond = c("BO", "BS", "CPS", "DE", "DT", "E1", "GB", "LO", "MI", "PB"),
           treatment = c("High", "Low"),
           tadpole = unique(formod$tadpole)[1])

testdata <- testdata %>%
  mutate(pred = predict(dev_mod_pond, newdata = testdata, allow.new.levels = T))

testdata %>%
  ggplot(aes(x = photo.date, y = pred, col = treatment))+
  geom_smooth(method = "lm", aes(group = interaction(pond, treatment)), lwd = 0.2)

# SPLIT the data into a list of data frames, each containing data from one pond at one treatment
l <- testdata %>%
  group_by(pond, treatment) %>%
  group_split() # split into list

# Function to dev rate (fit linear model and get slope) [this is what would go inside the for loop]
get_dev_rate <- function(df){
  mod <- lm(pred ~ photo.date, data = df) # make model
  coefs <- summary(mod)$coefficients %>% as.data.frame() # extract coefficients
  output <- data.frame(pond = df$pond[1], treatment = df$treatment[1], dev_rate = coefs$Estimate[2]) # put slope coefficient together with pond and treatment information for output
  return(output)
}

# APPLY the function to each of the split up data frames
dev_rates_list <- lapply(l, get_dev_rate) 

# COMBINE the list into a data frame
dev_rates <- rbindlist(dev_rates_list) 
head(dev_rates)

# Get long-term temperatures
longtemps <- read.csv("data/inputs/PTemps.predicted.2001_2018.csv", as.is = T) %>%
  select(Pond, PTemp) %>%
  rename(pond = Pond,
         ptemp = PTemp) %>%
  mutate(pond = case_when(pond == "CP" ~ "CPS",
                          pond == "LP" ~ "LO",
                          TRUE ~ pond))

save(longtemps, file = "data/outputs/longtemps.Rda")

devrates1 <- dev_rates %>% 
  left_join(longtemps, by = "pond")


head(devrates1)

devrates1 %>% ggplot(aes(x = ptemp, y = dev_rate))+
  geom_point(aes(col = treatment), size = 2)+
  scale_color_manual(values = c("orange", "dodgerblue3"))+
  geom_smooth(method = "lm", aes(col = treatment))
```


```{r}
load("data/outputs/avg_inc_temps.Rda")
incubator_dev_rates_bypond <- dev_rates %>%
  left_join(avg_inc_temps, by = "treatment")
save(incubator_dev_rates_bypond, file = "data/outputs/incubator_dev_rates_bypond.Rda")

incubator_dev_rates_bypond %>%
  ggplot(aes(x = avg_temp, y = dev_rate, col = treatment))+
  geom_smooth(aes(group = pond), method = "lm", col = "black", lty = "dashed", lwd = 0.5)+
  geom_point(size = 3, alpha = 0.7)+
  scale_color_manual(name = "Temperature treatment", values = c("orange", "dodgerblue3"))+
  theme_minimal()+
  theme(legend.position = "bottom")+
  ylab("Developmental rate (GS/day)")+
  xlab("Average incubator temperature")
```

