---
title: "figures_flipbook"
author: "Kaija Gahm"
date: "3/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load libraries
```{r}
source("libraries.R")
```

# Color, shape, and line type palettes
```{r}
col_hilo <- c("orange", "dodgerblue3")
col_hilowi <- c(col_hilo, "black")
shape_hilo <- c(20, 4)
linetype_hilo <- c("solid", "dashed")
linetype_hilowi <- c(linetype_hilo, "dotted")
```

# Incubators and pond temp loggers
```{r}
load("data/outputs/loggers_tojoin.Rda")
load("data/outputs/incs_long.Rda")

# Color
loggers_tojoin %>% ggplot(aes(x = date, y = mean_daily_temp_c, group = pond))+
  geom_line(alpha = 0.2)+
  geom_line(data = incs_long %>% filter(pond %in% c("high", "low")), 
            aes(x = date, y = temp_c, col = pond), lwd = 1.5)+
  theme_minimal()+
  xlab("Date")+
  ylab("Temperature (ºC)")+
  scale_color_manual(name = "Temperature treatment incubator", values = col_hilo, labels = c("High", "Low"))+
  theme(legend.position = "bottom")

# Black and white
loggers_tojoin %>% ggplot(aes(x = date, y = mean_daily_temp_c, group = pond))+
  geom_line(alpha = 0.2)+
  geom_line(data = incs_long %>% filter(pond %in% c("high", "low")), 
            aes(x = date, y = temp_c, lty = pond), lwd = 1)+
  theme_minimal()+
  xlab("Date")+
  ylab("Temperature (ºC)")+
  scale_linetype_manual(name = "Temperature treatment incubator", values = linetype_hilo, labels = c("High", "Low"))+
  guides(linetype = guide_legend(override.aes = list(lwd = 0.5)))+
  theme(legend.position = "bottom")

#~~~~~~~~~~~~~ Temp boxplots
load("data/outputs/fortempboxplots.Rda")

fortempboxplots %>% ggplot(aes(x = pond, y = mean_daily_temp_c))+
  geom_boxplot()+
  xlab("Pond")+
  ylab("Average daily temperature (ºC)")
```


# Speed by mass and temperature treatment
```{r}
load("data/outputs/dat3_nonas.Rda")

# Color
dat3_nonas %>% ggplot(aes(x = mass, y = sline_speed_mm_s, col = treatment))+
  geom_point()+
  geom_smooth(method = "lm", se = F)+
  ylab("Burst speed (mm/s)")+
  xlab("Mass (g)")+
  theme_minimal()+
  scale_color_manual(name = "Temperature treatment", values = col_hilo)+
  theme(legend.position = "bottom")

# Black and white
dat3_nonas %>% ggplot(aes(x = mass, y = sline_speed_mm_s))+
  geom_point(aes(shape = treatment))+
  geom_smooth(aes(lty = treatment), method = "lm", 
              col = "black", 
              lwd = 0.7,
              se = F)+
  ylab("Burst speed (mm/s)")+
  xlab("Mass (g)")+
  theme_minimal()+
  scale_shape_manual(values = shape_hilo)+
  scale_linetype_manual(values = linetype_hilo)+
  guides(shape = guide_legend(title = "", override.aes = list(alpha = 1)))+
  guides(lty = guide_legend(title = "Temperature \n treatment"))+
  theme(legend.position = "bottom")
```

# Development rate over time
```{r}
load("data/outputs/dev4.Rda")
load("data/outputs/dat3_nonas.Rda")
load("data/outputs/wild_nonas.Rda")

dev4 <- dev4 %>% filter(tadpole %in% c(dat3_nonas$tadpole, wild_nonas$tadpole))

# Color
dev4 %>% filter(!is.na(pond), !is.na(stage)) %>%
  ggplot(aes(x = photo.date, y = stage, col = treatment, group = tadpole))+
  #geom_point(size = 0.5)+
  geom_smooth(se = F, method = "lm", lwd = 0.05)+
  theme_minimal()+
  scale_color_manual(values = col_hilowi, name = "Temperature treatment")+
  guides(color = guide_legend(override.aes = list(size=5)))+
  xlab("Date")+
  ylab("Gosner stage")+
  theme(legend.position = "bottom")

# Color with points only for wild
dev4 %>% filter(!is.na(pond), !is.na(stage)) %>%
  filter(treatment %in% c("Low", "High")) %>%
  ggplot(aes(x = photo.date, y = stage, col = treatment, group = tadpole))+
  geom_smooth(se = F, method = "lm", lwd = 0.05)+
  geom_point(data = dev4 %>% filter(!is.na(pond), !is.na(stage), treatment == "Wild", photo.date > "2019-05-01"), 
             aes(x = photo.date, y = stage),
             alpha = 0.3)+
  theme_minimal()+
  scale_color_manual(values = col_hilowi, name = "Temperature treatment")+
  guides(color = guide_legend(override.aes = list(size = 5)))+
  xlab("Date")+
  ylab("Gosner stage")+
  theme(legend.position = "bottom")

# Color with points for all
finalpoints <- dev4 %>% # only the final photos for the high and low tadpoles.
  filter(!is.na(pond), !is.na(stage), treatment != "Wild") %>% 
  group_by(tadpole) %>%
  slice(n())

dev4 %>% filter(!is.na(pond), !is.na(stage)) %>%
  filter(treatment != "Wild") %>%
  ggplot(aes(x = photo.date, y = stage, col = treatment, group = tadpole))+
  geom_smooth(se = F, method = "lm", lwd = 0.05)+
  geom_point(data = dev4 %>% filter(!is.na(pond), !is.na(stage), treatment == "Wild", photo.date > "2019-05-01"), 
             aes(x = photo.date, y = stage),
             alpha = 0.3)+
  geom_point(data = finalpoints %>% filter(photo.date > "2019-05-01"),
             aes(x = photo.date, y = stage, col = treatment, group = tadpole),
             alpha = 0.3)+
  theme_minimal()+
  scale_color_manual(values = col_hilowi, name = "Temperature treatment")+
  guides(color = guide_legend(override.aes = list(size = 5)))+
  xlab("Date")+
  ylab("Gosner stage")+
  theme(legend.position = "bottom")

# Black and white
dev4 %>% filter(!is.na(pond), !is.na(stage)) %>%
  ggplot(aes(x = photo.date, y = stage, group = treatment))+
  #geom_point(size = 0.5)+
  geom_smooth(method = "lm", se = FALSE, aes(lty = treatment), col = "black")+
  scale_linetype_manual(values = linetype_hilowi)+
  theme_minimal()+
  xlab("Date")+
  ylab("Gosner stage")+
  guides(lty = guide_legend(override.aes = list(lwd = 0.5)))+
  guides(lty = guide_legend(title = "Temperature \n treatment"))+
  theme(legend.position = "bottom")

# Black and white with points only for wild
# I can't figure out how to do this one!


# Linear models
low <- lm(stage ~ photo.date, data = dev4 %>% filter(!is.na(pond), !is.na(stage), treatment == "Low"))
summary(low)
high <- lm(stage ~ photo.date, data = dev4 %>% filter(!is.na(pond), !is.na(stage), treatment == "High"))
summary(high) 
wild <- lm(stage ~ photo.date, data = dev4 %>% filter(!is.na(pond), !is.na(stage), treatment == "Wild"))
summary(wild)
```

# Plasticity by pond and clutch
```{r}
load("data/outputs/plasticity.Rda")

# Get order of ponds by mean plasticity
mean_pond_order <- plasticity %>% 
  group_by(pond) %>% 
  summarize(mean = mean(D, na.rm = T)) %>%
  arrange(-mean) %>%
  pull(pond) %>%
  as.character()

# Make plot
## Define helper functions
reorder_within <- function(x, by, within, fun = mean, sep = "___", ...) {
  new_x <- paste(x, within, sep = sep)
  stats::reorder(new_x, by, FUN = fun)
}

scale_x_reordered <- function(..., sep = "___") {
  reg <- paste0(sep, ".+$")
  ggplot2::scale_x_discrete(labels = function(x) gsub(reg, "", x), ...)
}

## make the plot
plasticity %>%
  mutate(pond = fct_relevel(pond, mean_pond_order)) %>%
  filter(!is.na(clutch_only)) %>%
  ggplot(aes(x = reorder_within(clutch_only, -D, pond, mean), y = D))+
  theme_minimal()+
  geom_rect(aes(fill = pond),
            xmin = -Inf, xmax = Inf,
            ymin = -Inf, ymax = Inf, 
            alpha = 0.3)+
  scale_fill_manual(values = rep(c("grey95", "grey78"), 5))+
  geom_boxplot()+
  scale_x_reordered()+
  facet_grid(cols = vars(pond), scales = "free_x")+
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    legend.position = "none"
  )+
  ylab("Scope of plastic response (low temp - high temp)")+
  xlab("Pond")
```

# Wild tadpoles: final model
```{r}
load("data/outputs/wild_nonas.Rda")
load("data/outputs/consensus_mod_wild.Rda")
# How do we plot this model properly?

wild_nonas %>% ggplot(aes(x = mass, y = sline_speed_mm_s))+
  geom_point(alpha = 0.4)+
  geom_smooth(method = "lm", col = "black")+
  xlab("Mass")+
  ylab("Burst speed (mm/s)")+
  theme_minimal()

# Should add the regular tadpole data along with wild and plot the two together.
```

# Comparison of high, low, and wild tadpoles in the same ponds
```{r}
load("data/outputs/indiv_means.Rda")

# Helper function to shift the legend
library(gtable)
library(cowplot)
library(grid)

shift_legend <- function(p){

  # check if p is a valid object
  if(!"gtable" %in% class(p)){
    if("ggplot" %in% class(p)){
      gp <- ggplotGrob(p) # convert to grob
    } else {
      message("This is neither a ggplot object nor a grob generated from ggplotGrob. Returning original plot.")
      return(p)
    }
  } else {
    gp <- p
  }

  # check for unfilled facet panels
  facet.panels <- grep("^panel", gp[["layout"]][["name"]])
  empty.facet.panels <- sapply(facet.panels, function(i) "zeroGrob" %in% class(gp[["grobs"]][[i]]))
  empty.facet.panels <- facet.panels[empty.facet.panels]
  if(length(empty.facet.panels) == 0){
    message("There are no unfilled facet panels to shift legend into. Returning original plot.")
    return(p)
  }

  # establish extent of unfilled facet panels (including any axis cells in between)
  empty.facet.panels <- gp[["layout"]][empty.facet.panels, ]
  empty.facet.panels <- list(min(empty.facet.panels[["t"]]), min(empty.facet.panels[["l"]]),
                             max(empty.facet.panels[["b"]]), max(empty.facet.panels[["r"]]))
  names(empty.facet.panels) <- c("t", "l", "b", "r")

  # extract legend & copy over to location of unfilled facet panels
  guide.grob <- which(gp[["layout"]][["name"]] == "guide-box")
  if(length(guide.grob) == 0){
    message("There is no legend present. Returning original plot.")
    return(p)
  }
  gp <- gtable_add_grob(x = gp,
                        grobs = gp[["grobs"]][[guide.grob]],
                        t = empty.facet.panels[["t"]],
                        l = empty.facet.panels[["l"]],
                        b = empty.facet.panels[["b"]],
                        r = empty.facet.panels[["r"]],
                        name = "new-guide-box")

  # squash the original guide box's row / column (whichever applicable)
  # & empty its cell
  guide.grob <- gp[["layout"]][guide.grob, ]
  if(guide.grob[["l"]] == guide.grob[["r"]]){
    gp <- gtable_squash_cols(gp, cols = guide.grob[["l"]])
  }
  if(guide.grob[["t"]] == guide.grob[["b"]]){
    gp <- gtable_squash_rows(gp, rows = guide.grob[["t"]])
  }
  gp <- gtable_remove_grobs(gp, "guide-box")

  return(gp)
}


# Color
pc <- indiv_means %>%
  ggplot(aes(x = mass, y = speed, col = treatment))+
  geom_smooth(method = "lm", alpha = 0.2)+
  facet_wrap(~pond)+
  theme_minimal()+
  scale_color_manual(name = "Temperature \n treatment", values = col_hilowi)+
  theme(legend.position = "bottom")+
  ylab("Burst speed (mm/s)")+
  xlab("Mass (g)")

grid.draw(shift_legend(pc)) # have to run this in the console for some reason.

# Black and white
pbw <- indiv_means %>%
  ggplot(aes(x = mass, y = speed))+
  geom_smooth(method = "lm", 
              col = "black", 
              alpha = 0.2,
              lwd = 0.7,
              aes(lty = treatment))+
  facet_wrap(~pond)+
  theme_minimal()+
  scale_linetype_manual(name = "Temperature \n treatment", values = linetype_hilowi)+
  theme(legend.position = "bottom")+
  xlab("Mass (g)")+
  ylab("Burst speed (mm/s)")

grid.draw(shift_legend(pbw)) # have to run this in the console for some reason

```

# Average size and stage of tadpoles at time of trials
```{r}
load("data/outputs/dat3_nonas.Rda")
dat3_nonas %>% select(tadpole) %>% distinct() %>% nrow()

dat3_nonas %>% 
  group_by(treatment) %>%
  summarize(avg_stage = mean(gs, na.rm = T),
            max_stage = max(gs, na.rm = T),
            min_stage = min(gs, na.rm = T),
            sd_stage = sd(gs, na.rm = T),
            avg_mass = mean(mass, na.rm = T),
            max_mass = max(mass, na.rm = T),
            min_mass = min(mass, na.rm = T),
            sd_mass = sd(mass, na.rm = T))

load("data/outputs/wild_nonas.Rda")
wild_nonas %>% 
  summarize(avg_stage = mean(gs, na.rm = T),
            max_stage = max(gs, na.rm = T),
            min_stage = min(gs, na.rm = T),
            sd_stage = sd(gs, na.rm = T),
            avg_mass = mean(mass, na.rm = T),
            max_mass = max(mass, na.rm = T),
            min_mass = min(mass, na.rm = T),
            sd_mass = sd(mass, na.rm = T))
```

# Model predictions (mass): lab-reared
```{r}
load("data/outputs/dat3_nonas.Rda")
load("data/outputs/consensus_mod.Rda")

# Function to calculate z scores
standardize <- function(vector){
  mu <- mean(vector)
  sd <- sd(vector)
  z_vector <- (vector-mu)/sd
  return(z_vector)
}

dat3_nonas <- dat3_nonas %>% mutate_at(.vars = c("PC1", "PC2", "mass", "size", "gs", "T_initial"), standardize)

preds <- dat3_nonas %>% 
  select(-c(sline_speed_mm_s, trial, predicted_speed, predicted_speed_withresids)) %>% 
  distinct() # create new df for predictions

preds$pred <- predict(object = consensus_mod, newdata = preds) + summary(consensus_mod)$sigma # make predictions

# Color
preds %>% ggplot(aes(x = mass, y = pred, col = treatment))+
  geom_point()+
  geom_smooth(method = "lm", 
              se = F)+
  ylab("Burst speed (mm/s)")+
  xlab("Mass (g)")+
  theme_minimal()+
  scale_color_manual(name = "Temperature treatment", values = col_hilo)+
  theme(legend.position = "bottom")

# Black and white
preds %>% ggplot(aes(x = mass, y = pred))+
  geom_point(aes(shape = treatment))+
  geom_smooth(aes(lty = treatment), 
              method = "lm", 
              col = "black", 
              lwd = 0.7,
              se = F)+
  ylab("Burst speed (mm/s)")+
  xlab("Mass (g)")+
  theme_minimal()+
  scale_shape_manual(values = shape_hilo)+
  scale_linetype_manual(values = linetype_hilo)+
  guides(shape = guide_legend(title = "", override.aes = list(alpha = 1)))+
  guides(lty = guide_legend(title = "Temperature \n treatment"))+
  theme(legend.position = "bottom")
```

# Model predictions (gs): lab-reared
```{r}
# Color
preds %>% ggplot(aes(x = jitter(gs), y = pred, col = treatment))+
  geom_point()+
  geom_smooth(method = "lm", 
              se = F)+
  ylab("Burst speed (mm/s)")+
  xlab("Gosner stage")+
  theme_minimal()+
  scale_color_manual(name = "Temperature treatment", values = col_hilo)+
  theme(legend.position = "bottom")

# Black and white
preds %>% ggplot(aes(x = jitter(gs), y = pred))+
  geom_point(aes(shape = treatment))+
  geom_smooth(aes(lty = treatment), 
              method = "lm", 
              col = "black", 
              lwd = 0.7,
              se = F)+
  ylab("Burst speed (mm/s)")+
  xlab("Gosner stage")+
  theme_minimal()+
  scale_shape_manual(values = shape_hilo)+
  scale_linetype_manual(values = linetype_hilo)+
  guides(shape = guide_legend(title = "", override.aes = list(alpha = 1)))+
  guides(lty = guide_legend(title = "Temperature \n treatment"))+
  theme(legend.position = "bottom")

# Residual of mass on stage
mass_stage_mod <- lm(mass ~ gs, data = preds)

preds$mass_stage_resid <- resid(mass_stage_mod)

# Color
preds %>% ggplot(aes(x = mass_stage_resid, y = pred, col = treatment))+
  geom_point()+
  geom_smooth(method = "lm", 
              se = F)+
  ylab("Burst speed (mm/s)")+
  xlab("Residual of mass at stage")+
  theme_minimal()+
  scale_color_manual(name = "Temperature treatment", values = col_hilo)+
  theme(legend.position = "bottom")

# Black and white
preds %>% ggplot(aes(x = mass_stage_resid, y = pred))+
  geom_point(aes(shape = treatment))+
  geom_smooth(aes(lty = treatment), 
              method = "lm", 
              col = "black", 
              lwd = 0.7,
              se = F)+
  ylab("Burst speed (mm/s)")+
  xlab("Residual of mass at stage")+
  theme_minimal()+
  scale_shape_manual(values = shape_hilo)+
  scale_linetype_manual(values = linetype_hilo)+
  guides(shape = guide_legend(title = "", override.aes = list(alpha = 1)))+
  guides(lty = guide_legend(title = "Temperature \n treatment"))+
  theme(legend.position = "bottom")
```

# Model predictions: wild
```{r}
load("data/outputs/wild_nonas.Rda")
head(wild_nonas)
load("data/outputs/consensus_mod_wild.Rda")

standardize <- function(vector){
  mu <- mean(vector)
  sd <- sd(vector)
  z_vector <- (vector-mu)/sd
  return(z_vector)
}

wild_nonas <- wild_nonas %>% mutate_at(.vars = c("PC1", "PC2", "mass", "size", "gs", "T_initial"), standardize)

wild_preds <- wild_nonas %>% 
  select(-c(sline_speed_mm_s, trial)) %>% 
  distinct() %>% # create new df for predictions
  select(mass, tadpole)

wild_preds$pred <- predict(object = consensus_mod_wild, newdata = wild_preds, re.form = NULL) + summary(consensus_mod_wild)$sigma # make predictions

wild_preds %>% ggplot(aes(x = mass, y = pred))+
  geom_point(alpha = 0.4)+
  geom_smooth(method = "lm", 
              col = "black",
              se = F)+
  xlab("Mass")+
  ylab("Burst speed (mm/s)")+
  theme_minimal()
```

# Number of trials
```{r}
load("data/outputs/dat3_nonas.Rda")
dat3_nonas %>% group_by(tadpole) %>% summarize(n = n()) %$% mean(n, na.rm = T) # average number of trials
dat3_nonas %>% pull(tadpole) %>% unique() %>% length() # number of lab-reared tadpoles
dat3_nonas %>% select(treatment, tadpole) %>% distinct() %$% table(treatment)


load("data/outputs/wild_nonas.Rda")
wild_nonas %>% group_by(tadpole) %>% summarize(n = n()) %$% mean(n, na.rm = T) # average number of trials
wild_nonas %>% pull(tadpole) %>% unique() %>% length() # number of wild tadpoles



wild_nonas %>% select(tadpole, trial) %>% rbind(dat3_nonas %>% select(tadpole, trial)) %>% group_by(tadpole) %>%
  summarize(n = n()) %$% mean(n, na.rm = T) # average number of trials overall

nrow(dat3_nonas) + nrow(wild_nonas)

```

# MAIN FIGURE: burst speed/mass, lab and wild
```{r}
load("data/outputs/dat3_nonas.Rda")
load("data/outputs/consensus_mod.Rda")

load("data/outputs/consensus_mod_wild.Rda")

# Standardize variables
# Function to calculate z scores
standardize <- function(vector){
  mu <- mean(vector)
  sd <- sd(vector)
  z_vector <- (vector-mu)/sd
  return(z_vector)
}

dat3_nonas <- dat3_nonas %>% mutate_at(.vars = c("PC1", "PC2", "mass", "size", "gs", "T_initial"), standardize)

# Get model predictions
preds <- dat3_nonas %>% 
  select(-c(sline_speed_mm_s, trial, predicted_speed, predicted_speed_withresids)) %>% 
  distinct() # create new df for predictions

preds$pred <- predict(object = consensus_mod, newdata = preds) + summary(consensus_mod)$sigma # make predictions

summary(consensus_mod)$sigma
summary(consensus_mod_wild)$sigma

# Color figure
preds %>% ggplot(aes(x = mass, y = pred, col = treatment))+
  geom_point()+
  geom_smooth(method = "lm", 
              se = F)+
  ylab("Burst speed (mm/s)")+
  xlab("Mass (g)")+
  theme_minimal()+
  scale_color_manual(name = "Temperature treatment", values = col_hilo)+
  theme(legend.position = "bottom")+
  geom_abline(intercept = 59.876, 
              slope = 4.276, 
              lwd = 1)
```


