---
title: "figures_flipbook"
author: "Kaija Gahm"
date: "3/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load libraries
```{r}
source("libraries.R")
```

# Color, shape, and line type palettes
```{r}
col_hilo <- c("orange", "dodgerblue3")
col_hilowi <- c(col_hilo, "black")
shape_hilo <- c(20, 4)
linetype_hilo <- c("solid", "dashed")
linetype_hilowi <- c(linetype_hilo, "dotted")
```

# Incubators and pond temp loggers
```{r}
load("data/outputs/loggers_cropped.Rda")
load("data/outputs/incs_long_cropped.Rda")

# Color
loggers_cropped %>% ggplot(aes(x = date, y = mean_daily_temp_c, group = pond))+
  geom_line(alpha = 0.2)+
  geom_line(data = incs_long_cropped, 
            aes(x = date, y = temp_c, col = pond), lwd = 1.5)+
  theme_minimal()+
  xlab("Date (2019)")+
  ylab("Temperature (ºC)")+
  scale_color_manual(name = "Temperature treatment incubator", values = col_hilo, labels = c("High", "Low"))+
  theme(legend.position = "bottom")

# Black and white
loggers_cropped %>% ggplot(aes(x = date, y = mean_daily_temp_c, group = pond))+
  geom_line(alpha = 0.2)+
  geom_line(data = incs_long_cropped, 
            aes(x = date, y = temp_c, lty = pond), lwd = 1)+
  theme_minimal()+
  xlab("Date (2019)")+
  ylab("Temperature (ºC)")+
  scale_linetype_manual(name = "Temperature treatment incubator", values = linetype_hilo, labels = c("High", "Low"))+
  guides(linetype = guide_legend(override.aes = list(lwd = 0.5)))+
  theme(legend.position = "bottom")

#~~~~~~~~~~~~~ Temp boxplots
load("data/outputs/temps_cropped.Rda")

temps_cropped %>%
  filter(!is.na(mean_daily_temp_c)) %>%
  ggplot(aes(x = fct_reorder(pond, -mean_daily_temp_c), 
             y = mean_daily_temp_c))+
  geom_boxplot()+
  xlab("Pond")+
  ylab("Average daily temperature")
```


# Speed by mass and temperature treatment
```{r}
load("data/outputs/dat3_nonas.Rda")

# Color
dat3_nonas %>% ggplot(aes(x = mass, y = sline_speed_mm_s, col = treatment))+
  geom_point()+
  geom_smooth(method = "lm", se = F)+
  ylab("Burst speed (mm/s)")+
  xlab("Mass (g)")+
  theme_minimal()+
  scale_color_manual(name = "Temperature treatment", values = col_hilo)+
  theme(legend.position = "bottom")

# Black and white
dat3_nonas %>% ggplot(aes(x = mass, y = sline_speed_mm_s))+
  geom_point(aes(shape = treatment))+
  geom_smooth(aes(lty = treatment), method = "lm", 
              col = "black", 
              lwd = 0.7,
              se = F)+
  ylab("Burst speed (mm/s)")+
  xlab("Mass (g)")+
  theme_minimal()+
  scale_shape_manual(values = shape_hilo)+
  scale_linetype_manual(values = linetype_hilo)+
  guides(shape = guide_legend(title = "", override.aes = list(alpha = 1)))+
  guides(lty = guide_legend(title = "Temperature \n treatment"))+
  theme(legend.position = "bottom")
```

# Development rate over time
```{r}
load("data/outputs/dev4.Rda")
load("data/outputs/dat3_nonas.Rda")
load("data/outputs/wild_nonas.Rda")

dev4 <- dev4 %>% filter(tadpole %in% c(dat3_nonas$tadpole, wild_nonas$tadpole))

# Color
dev4 %>% filter(!is.na(pond), !is.na(stage)) %>%
  ggplot(aes(x = photo.date, y = stage, col = treatment, group = tadpole))+
  #geom_point(size = 0.5)+
  geom_smooth(se = F, method = "lm", lwd = 0.05)+
  theme_minimal()+
  scale_color_manual(values = col_hilowi, name = "Temperature treatment")+
  guides(color = guide_legend(override.aes = list(size=5)))+
  xlab("Date")+
  ylab("Gosner stage")+
  theme(legend.position = "bottom")

# Color with points only for wild
dev4 %>% filter(!is.na(pond), !is.na(stage)) %>%
  filter(treatment %in% c("Low", "High")) %>%
  ggplot(aes(x = photo.date, y = stage, col = treatment, group = tadpole))+
  geom_smooth(se = F, method = "lm", lwd = 0.05)+
  geom_point(data = dev4 %>% filter(!is.na(pond), !is.na(stage), treatment == "Wild", photo.date > "2019-05-01"), 
             aes(x = photo.date, y = stage),
             alpha = 0.3)+
  theme_minimal()+
  scale_color_manual(values = col_hilowi, name = "Temperature treatment")+
  guides(color = guide_legend(override.aes = list(size = 5)))+
  xlab("Date")+
  ylab("Gosner stage")+
  theme(legend.position = "bottom")

# Color with points for all
finalpoints <- dev4 %>% # only the final photos for the high and low tadpoles.
  filter(!is.na(pond), !is.na(stage), treatment != "Wild") %>% 
  group_by(tadpole) %>%
  slice(n())

dev4 %>% filter(!is.na(pond), !is.na(stage)) %>%
  filter(treatment != "Wild") %>%
  ggplot(aes(x = photo.date, y = stage, col = treatment, group = tadpole))+
  geom_smooth(se = F, method = "lm", lwd = 0.05)+
  geom_point(data = dev4 %>% filter(!is.na(pond), !is.na(stage), treatment == "Wild", photo.date > "2019-05-01"), 
             aes(x = photo.date, y = stage),
             alpha = 0.3)+
  geom_point(data = finalpoints %>% filter(photo.date > "2019-05-01"),
             aes(x = photo.date, y = stage, col = treatment, group = tadpole),
             alpha = 0.3)+
  theme_minimal()+
  scale_color_manual(values = col_hilowi, name = "Temperature treatment")+
  guides(color = guide_legend(override.aes = list(size = 5)))+
  xlab("Date")+
  ylab("Gosner stage")+
  theme(legend.position = "bottom")

# Black and white
dev4 %>% filter(!is.na(pond), !is.na(stage)) %>%
  ggplot(aes(x = photo.date, y = stage, group = treatment))+
  #geom_point(size = 0.5)+
  geom_smooth(method = "lm", se = FALSE, aes(lty = treatment), col = "black")+
  scale_linetype_manual(values = linetype_hilowi)+
  theme_minimal()+
  xlab("Date")+
  ylab("Gosner stage")+
  guides(lty = guide_legend(override.aes = list(lwd = 0.5)))+
  guides(lty = guide_legend(title = "Temperature \n treatment"))+
  theme(legend.position = "bottom")

# Black and white with points only for wild
# I can't figure out how to do this one!
```

# Plasticity by pond and clutch
```{r}
load("data/outputs/plasticity.Rda")

# Get order of ponds by mean plasticity
mean_pond_order <- plasticity %>% 
  group_by(pond) %>% 
  summarize(mean = mean(D, na.rm = T)) %>%
  arrange(-mean) %>%
  pull(pond) %>%
  as.character()

# Make plot
## Source helper functions
source("plot_helper_funs.R")

## make the plot
plasticity %>%
  mutate(pond = fct_relevel(pond, mean_pond_order)) %>%
  filter(!is.na(clutch_only)) %>%
  ggplot(aes(x = reorder_within(clutch_only, -D, pond, mean), y = D))+
  theme_minimal()+
  geom_rect(aes(fill = pond),
            xmin = -Inf, xmax = Inf,
            ymin = -Inf, ymax = Inf, 
            alpha = 0.3)+
  scale_fill_manual(values = rep(c("white", "grey90"), 5))+
  geom_hline(aes(yintercept = 0), color = "black")+
  geom_boxplot()+
  scale_x_reordered()+
  facet_grid(cols = vars(pond), scales = "free_x")+
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    legend.position = "none"
  )+
  ylab("Scope of plastic response (low - high)")+
  xlab("Pond")

sample_sizes <- plasticity %>% 
  group_by(pond, clutch) %>%
  summarize(n = n())

summary(sample_sizes$n)
```

# Wild tadpoles: final model
```{r}
load("data/outputs/wild_nonas.Rda")
load("data/outputs/consensus_mod_wild.Rda")
# How do we plot this model properly?

wild_nonas %>% ggplot(aes(x = mass, y = sline_speed_mm_s))+
  geom_point(alpha = 0.4)+
  geom_smooth(method = "lm", col = "black", se = F)+
  xlab("Mass")+
  ylab("Burst speed (mm/s)")+
  theme_minimal()
```

# Average size and stage of tadpoles at time of trials
```{r}
load("data/outputs/dat3_nonas.Rda")
dat3_nonas %>% select(tadpole) %>% distinct() %>% nrow()

dat3_nonas %>% 
  group_by(treatment) %>%
  summarize(avg_stage = mean(gs, na.rm = T),
            max_stage = max(gs, na.rm = T),
            min_stage = min(gs, na.rm = T),
            sd_stage = sd(gs, na.rm = T),
            avg_mass = mean(mass, na.rm = T),
            max_mass = max(mass, na.rm = T),
            min_mass = min(mass, na.rm = T),
            sd_mass = sd(mass, na.rm = T))

load("data/outputs/wild_nonas.Rda")
wild_nonas %>% 
  summarize(avg_stage = mean(gs, na.rm = T),
            max_stage = max(gs, na.rm = T),
            min_stage = min(gs, na.rm = T),
            sd_stage = sd(gs, na.rm = T),
            avg_mass = mean(mass, na.rm = T),
            max_mass = max(mass, na.rm = T),
            min_mass = min(mass, na.rm = T),
            sd_mass = sd(mass, na.rm = T))
```

# Number of trials
```{r}
load("data/outputs/dat3_nonas.Rda")
dat3_nonas %>% group_by(tadpole) %>% summarize(n = n()) %$% mean(n, na.rm = T) # average number of trials
dat3_nonas %>% pull(tadpole) %>% unique() %>% length() # number of lab-reared tadpoles
dat3_nonas %>% select(treatment, tadpole) %>% distinct() %$% table(treatment)


load("data/outputs/wild_nonas.Rda")
wild_nonas %>% group_by(tadpole) %>% summarize(n = n()) %$% mean(n, na.rm = T) # average number of trials
wild_nonas %>% pull(tadpole) %>% unique() %>% length() # number of wild tadpoles



wild_nonas %>% select(tadpole, trial) %>% rbind(dat3_nonas %>% select(tadpole, trial)) %>% group_by(tadpole) %>%
  summarize(n = n()) %$% mean(n, na.rm = T) # average number of trials overall

nrow(dat3_nonas) + nrow(wild_nonas)

```

# Allocation
```{r}
load("data/outputs/dat3_nonas.Rda")
load("data/outputs/wild_nonas.Rda")
plot.alloc <- dat3_nonas %>% ggplot(aes(x = jitter(gs, amount = 0.1), y = mass, col = treatment))+ 
  geom_point(alpha = 0.5)+
  geom_smooth(method = "lm")+
  geom_point(data = wild_nonas, color = "black", alpha = 0.5)+ # add points and line for wild tadpoles
  geom_smooth(data = wild_nonas, method = "lm", col = "black")+
  scale_color_manual(name = "Temperature treatment", values = col_hilo)+
  theme_minimal()+
  theme(legend.position = "bottom")+
  ylab("Mass (g)")+
  xlab("Gosner stage")
plot.alloc
```

# Partial regression plots
## Data
### Load
```{r}
load("data/outputs/newlab.gs.Rda")
load("data/outputs/newwld.gs.Rda")
load("data/outputs/newlab.sas.Rda")
load("data/outputs/newwld.sas.Rda")
```

### Bind
```{r}
newgs <- bind_rows(newlab.gs, newwld.gs) %>%
  mutate(treatment = case_when(is.na(treatment) ~ "Wild",
                               TRUE ~ as.character(treatment))) %>%
  mutate(treatment = factor(treatment))

newsas <- bind_rows(newlab.sas, newwld.sas) %>%
  mutate(treatment = case_when(is.na(treatment) ~ "Wild",
                               TRUE ~ as.character(treatment))) %>%
  mutate(treatment = factor(treatment))
```

### Save
```{r}
save(newgs, file = "data/outputs/newgs.Rda")
save(newsas, file = "data/outputs/newsas.Rda")
```


## Points
```{r}
# Gosner stage (points)
plot.gs <- newgs %>%
  ggplot(aes(x = jitter(gs, amount = 0.1), y = fit.re, col = treatment))+
  geom_point(alpha = 0.5)+
  geom_smooth(aes(y = fit.nore), method = "lm", se = F)+
  scale_color_manual(values = col_hilowi)+
  ylab("Predicted speed (mm/s)")+
  xlab("")+
  theme_minimal()+
  theme(legend.position = "none")

# Mass at stage (points)
plot.sas <- newsas %>%
  ggplot(aes(x = mass_at_stage, y = fit.re, col = treatment))+
  geom_point(alpha = 0.5)+
  geom_smooth(aes(y = fit.nore), method = "lm", se = F)+
  scale_color_manual(values = col_hilowi)+
  ylab("")+
  xlab("")+
  theme_minimal()+
  theme(legend.position = "none")
```

## Lines
```{r}
# Gosner stage (lines)
plot.gs.re <- newgs %>%
  ggplot(aes(x = gs, y = fit.re, col = treatment))+
  geom_smooth(aes(y = fit.re, group = interaction(pond, treatment)),
              method = "lm", se = F, size = 0.5)+
  geom_smooth(aes(y = fit.nore),
              method = "lm", se = F, size = 2)+
  scale_color_manual(values = col_hilowi)+
  ylab("Predicted speed (mm/s)")+
  xlab("Gosner stage")+
  theme_minimal()+
  theme(legend.position = "bottom")
plot.gs.re

# Mass at stage (lines)
plot.sas.re <- newsas %>%
  ggplot(aes(x = mass_at_stage, y = fit.re, col = treatment))+
  geom_smooth(aes(y = fit.re, group = interaction(pond, treatment)),
              method = "lm", se = F, size = 0.5)+
  geom_smooth(aes(y = fit.nore),
              method = "lm", se = F, size = 2)+
  scale_color_manual(name = "Temperature \n treatment", values = col_hilowi) +
  ylab("")+
  xlab("Mass at stage")+
  theme_minimal()+
  theme(legend.position = "none")
plot.sas.re
```

## Cowplots
```{r}
cowplot::plot_grid(plot.gs, plot.sas, plot.gs.re, plot.sas.re, nrow = 2)
cowplot::plot_grid(plot.gs.re, plot.sas.re, align = "h")
```

## By pond
```{r}
source("plot_helper_funs.R")
pondplot <- newgs %>%
  ggplot(aes(x = gs, y = fit.re, col = treatment))+ # use fit.re here bc it's different for each pond
  geom_smooth(method = "lm", se = F)+
  scale_color_manual(name = "Temperature treatment", values = col_hilowi)+
  theme_minimal()+
  facet_wrap(~pond)+
  ylab("Predicted speed (mm/s)")+
  xlab("Gosner stage")

grid.draw(shift_legend(pondplot)) # have to run this in the console for some reason
```

# Reaction norms
```{r}
# Lab vs. wild
load("data/outputs/dat3_nonas.Rda")
load("data/outputs/wild_nonas.Rda")
load("data/outputs/consensus_mod.Rda")
load("data/outputs/consensus_mod_wild.Rda")

head(dat3_nonas)

lab_pred <- dat3_nonas %>%
  mutate(predicted_speed = predict(consensus_mod, newdata = dat3_nonas, re.form = NULL))

wild_pred <- wild_nonas %>%
  mutate(predicted_speed = predict(consensus_mod_wild, newdata = wild_nonas, re.form = NULL))

lab <- lab_pred %>%
  group_by(treatment, pond) %>%
  mutate(meanspeed = mean(predicted_speed),
         sd = sd(predicted_speed)) %>%
  select(pond, meanspeed, sd, avg_temp) %>%
  distinct()

wild <- wild_pred %>%
  group_by(pond) %>%
  mutate(meanspeed = mean(predicted_speed),
         sd = sd(predicted_speed)) %>%
  select(pond, meanspeed, sd, avg_pond_temp) %>%
  distinct()


lab %>%
  ggplot(aes(x = avg_temp, y = meanspeed))+
  geom_smooth(data = wild, aes(x = avg_pond_temp, y = meanspeed),
              method = "lm", se = T, col = "black", lwd = 0.5)+
  geom_point(data = wild, aes(x = avg_pond_temp, y = meanspeed), size = 2)+
  scale_color_manual(name = "Temperature treatment", values = col_hilo)+
  geom_point(aes(col = treatment), size = 3, alpha = 0.7)+
  geom_smooth(method = "lm", se = F, aes(group = pond),
              col = "black", lwd = 0.3, lty = "dashed")+
  theme_minimal()+
  ylab("Burst speed (mm/s)")+
  xlab("Average temperature (ºC)")+
  theme(legend.position = "bottom")+
  guides(col = guide_legend(override.aes = list(alpha = 1)))

load("data/outputs/pond_avg_temps.Rda")
load("data/outputs/dat3_nonas.Rda")


lab %>%
  left_join(pond_avg_temps, by = "pond") %>%
  ggplot(aes(x = avg_pond_temp, y = meanspeed, col = treatment))+
  geom_point()+
  geom_smooth(method = "lm")+
  scale_color_manual(name = "Temperature treatment", values = col_hilo)+
  facet_wrap(~treatment)+
  theme_minimal()+
  ylab("Mean predicted burst speed")+
  xlab("Mean pond temperature")+
  theme(legend.position = "bottom")

mod <- lm(meanspeed ~ avg_pond_temp + treatment, data = lab %>% left_join(pond_avg_temps, by = "pond"))
summary(mod)


# Development vs. burst speed

# Developmental rate
load("data/outputs/incubator_dev_rates_bypond.Rda")
dev.plot <- incubator_dev_rates_bypond %>%
  ggplot(aes(x = avg_temp, y = dev_rate, col = treatment))+
  geom_smooth(aes(group = pond), method = "lm", col = "black", lty = "dashed", lwd = 0.5)+
  geom_point(size = 3, alpha = 0.7)+
  scale_color_manual(name = "Temperature treatment", values = c("orange", "dodgerblue3"))+
  theme_minimal()+
  theme(legend.position = "none")+
  ylab("Developmental rate (GS/day)")+
  xlab("")+
  guides(col = guide_legend(override.aes = list(alpha = 1)))

# Burst speed
burstspeed.plot <- lab %>%
  ggplot(aes(x = avg_temp, y = meanspeed))+
  geom_point(aes(col = treatment), size = 3, alpha = 0.7)+
  geom_smooth(method = "lm", se = F, aes(group = pond),
              col = "black", lwd = 0.3, lty = "dashed")+
  scale_color_manual(name = "Temperature treatment", values = col_hilo)+
  theme_minimal()+
  ylab("Burst speed (mm/s)")+
  xlab("Average temperature (ºC)")+
  theme(legend.position = "bottom")+
  guides(col = guide_legend(override.aes = list(alpha = 1)))

cowplot::plot_grid(dev.plot, burstspeed.plot, nrow = 2, align = "h")
```


