---
title: "JEZ_Analysis"
author: "Kaija Gahm"
date: "4/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries
```{r}
source("libraries.R")
```

# Load data
```{r}
# Lab and wild
load("data/outputs/dat3_nonas.Rda")
load("data/outputs/wild_nonas.Rda")

# Oviposition dates
load("data/outputs/wild_dates.Rda")
head(wild_dates)
```

# Calculate developmental rates (stages/day)
## Wild
### Calculate
```{r}
wild <- wild_nonas %>%
  mutate(devrate = (gs-1)/as.numeric((burst_date-ovip_date)))
```

## Lab
### Bind oviposition dates
```{r}
lab <- wild_dates %>%
  select(pond, ovip_date) %>%
  distinct() %>%
  right_join(dat3_nonas, by = "pond")
```
### Calculate
```{r}
lab <- lab %>%
  mutate(devrate = (gs-1)/as.numeric((burst_date-ovip_date)))

head(lab)
```

# LAB
## Do dev rates differ between Tx?
### ANOVA
```{r}
mod <- lm(devrate ~ treatment, data = lab)
summary(mod)
```

## Does performance differ between Tx?
### Mixed model
```{r}
load("data/outputs/dat3_nonas.Rda")

# Unstandardized
performance_mod_lab <- lmer(sline_speed_mm_s ~ mass_at_stage*treatment + gs*treatment + PC2 + T_initial + (1|pond/clutch/tadpole), data = dat3_nonas)
summary(performance_mod_lab)

# Standardized
load("data/outputs/dat3_nonas_std.Rda")

performance_mod_lab_std <- lmer(sline_speed_mm_s ~ mass_at_stage*treatment + gs*treatment + PC2 + T_initial + (1|pond/clutch/tadpole), data = dat3_nonas_std)
summary(performance_mod_lab_std)
```

### Partial regression plots
#### Mean-constant data (GS)
Hold all variables constant at their mean values for lab tadpoles except for Gosner stage. Mass_at_stage will be held to 0, since those values are residuals.

We'll use unstandardized data for this.
```{r}
# Make new datasets
newlab.gs <- dat3_nonas %>% # make new means/medians dataset (lab)
  select(tadpole, clutch, mass_at_stage, treatment, pond, T_initial, gs, PC2) %>%
  mutate(T_initial = mean(T_initial),
         mass_at_stage = mean(mass_at_stage),
         PC2 = mean(PC2))

# Add predictions
newlab.gs <- newlab.gs %>%
  mutate(fit.nore = predict(performance_mod_lab, newdata = newlab.gs, re.form = ~0),
         fit.re = predict(performance_mod_lab, newdata = newlab.gs, re.form = NULL))

newlab.gs <- newlab.gs %>%
  distinct()
```

##### Plot (GS)
```{r}
plot.gs <- newlab.gs %>%
  ggplot(aes(x = jitter(gs, amount = 0.1), y = fit.re, col = treatment))+
  geom_point(alpha = 0.75)+
  geom_smooth(aes(y = fit.nore),
              method = "lm", se = F)+
  scale_color_manual(name = "Temperature treatment",
                     values = c("orange", "dodgerblue3"))+
  ylab("Predicted speed (mm/s)")+
  xlab("Gosner stage")+
  theme_minimal()+
  theme(legend.position = "bottom")
plot.gs
```

#### Mean-constant data (mass at stage)
```{r}
# Make new datasets
newlab.mas <- dat3_nonas %>%
  select(tadpole, clutch, mass_at_stage, treatment, pond, T_initial, gs, PC2) %>%
  mutate(T_initial = mean(T_initial),
         gs = median(gs),
         PC2 = mean(PC2))

# Add predictions
newlab.mas <- newlab.mas %>%
  mutate(fit.nore = predict(performance_mod_lab, newdata = newlab.mas, re.form = ~0),
         fit.re = predict(performance_mod_lab, newdata = newlab.mas, re.form = NULL))

newlab.mas <- newlab.mas %>%
  distinct()
```

##### Plot (mass at stage)
```{r}
plot.mas <- newlab.mas %>%
  ggplot(aes(x = mass_at_stage, y = fit.re, col = treatment))+
  geom_point(alpha = 0.75) +
  geom_smooth(aes(y = fit.nore),
              method = "lm", se = F)+
  scale_color_manual(name = "Temperature treatment",
                     values = c("orange", "dodgerblue3"))+
  ylab("Predicted speed (mm/s)")+
  xlab("Mass at stage")+
  theme_minimal()+
  theme(legend.position = "bottom")
plot.mas
```

##### Plot (both)
```{r}
cowplot::plot_grid(plot.gs, plot.mas)
```

#### Without points (RE by pond)
I think the plots above are a little misleading because they don't show clear how much variaition there is in the wild compared to the lab tadpoles. The lab animals have really tight distributions, but the wild animals are all over the place and vary a ton just due to the pond that they were from.

Below I've dropped the points and instead am plotting the random effects by pond to show the variation. -Andis
##### GS
```{r}
plot.gs.re <- newlab.gs %>%
  ggplot(aes(x = gs, y = fit.re, col = treatment))+
  geom_smooth(aes(y = fit.re, group = interaction(pond, treatment)),
              method = "lm", se = F, size = 0.5)+
  geom_smooth(aes(y = fit.nore),
              method = "lm", se = F, size = 2)+
  scale_color_manual(name = "Temperature treatment",
                     values = c("orange", "dodgerblue3"))+
  ylab("Predicted speed (mm/s)")+
  xlab("Gosner stage")+
  theme_minimal()+
  theme(legend.position = "bottom")
plot.gs.re
```

##### Mass at stage
```{r}
plot.mas.re <- newlab.mas %>%
  ggplot(aes(x = mass_at_stage, y = fit.re, col = treatment))+
  geom_smooth(aes(y = fit.re, group = interaction(pond, treatment)),
              method = "lm", se = F, size = 0.5)+
  geom_smooth(aes(y = fit.nore),
              method = "lm", se = F, size = 2)+
  scale_color_manual(name = "Temperature treatment",
                     values = c("orange", "dodgerblue3"))+
  ylab("Predicted speed (mm/s)")+
  xlab("Mass at stage")+
  theme_minimal()+
  theme(legend.position = "bottom")
plot.mas.re
```

##### Both
```{r}
cowplot::plot_grid(plot.gs.re, plot.mas.re)
```

## Tradeoff? (Dev/performance)
### Make data
```{r}
## Create a new dataset with all of the main effect variables set to the mean BUT maintain treatment values.
new.lab <- lab %>% mutate(
  mass_at_stage = mean(mass_at_stage, na.rm = T),
  gs = median(gs, na.rm = T),
  PC2 = mean(PC2, na.rm = T),
  T_initial = mean(T_initial, na.rm = T),
  treatment = treatment
  )

# Fit the predicted speed for each individual, since we held the fixed effects constant, these are essentially just the conditional modes of the random effects. We also want to maintain our individual level variation, so we include the individual residuals from the model.
lab_tradeoff <- lab %>% mutate(
  fit.re = predict(performance_mod_lab, newdata = new.lab, re.form = NULL),
  resid = summary(performance_mod_lab)$resid,
  fit = fit.re + resid) %>%
  select(pond, clutch, tadpole, treatment, devrate, fit)

head(lab_tradeoff)
```

### Plot
```{r}
lab_tradeoff_plot <- lab_tradeoff %>%
  group_by(treatment, pond, clutch, tadpole) %>% 
  summarize(devrate = mean(devrate), fit = mean(fit)) %>%
  group_by(treatment, pond, clutch) %>%
  summarize(devrate = mean(devrate), fit = mean(fit)) %>%
  group_by(treatment, pond) %>%
  summarize(devrate = mean(devrate), fit = mean(fit)) %>%
  ggplot(aes(x = devrate, y = fit, col = treatment))+
  geom_point(size = 3, alpha = 0.6)+
  geom_line(aes(group = pond), lty = "dashed", lwd = 0.5, col = "black")+
  scale_color_manual(name = "Temperature treatment", values = c("orange", "dodgerblue3"))+
  theme_minimal()+
  theme(legend.position = "bottom")+
  ylab("Predicted burst speed (mm/s)")+
  xlab("Developmental rate (GS/day)")
lab_tradeoff_plot
```

# Wild
## How do wild dev rates compare?
### Prepare data
```{r}
wilddev <- wild %>% 
  select(devrate) %>%
  mutate(treatment = "Wild")

labdev <- lab %>%
  select(devrate, treatment)

dev <- bind_rows(wilddev, labdev)
```

### ANOVA
```{r}
mod <- lm(devrate ~ treatment, data = dev)
summary(mod)
```

## Do dev rates differ by pond?
### Model of devrate with pond as a factor
```{r}
# Are there differences in developmental rate by pond?
devmod <- lm(devrate ~ pond, data = wild)
summary(devmod)
visreg(devmod)
```

## Does performance differ by pond?
### Model of burst speed  with pond as a factor, including covariates
```{r}
# Unstandardized
performance_mod_wild <- lmer(sline_speed_mm_s ~ mass_at_stage * factor(pond) + gs * factor(pond) + PC2 + T_initial + (1|tadpole), data = wild_nonas)
summary(performance_mod_wild)

# Standardized
load("data/outputs/wild_nonas_std.Rda")

performance_mod_wild_std <- lmer(sline_speed_mm_s ~ mass_at_stage * factor(pond) + gs * factor(pond) + PC2 + T_initial + (1|tadpole), data = wild_nonas_std)
summary(performance_mod_wild_std)

# Showing differences by pond
visreg(performance_mod_wild_std, xvar = "pond", gg = T)+
  theme_minimal()+
  ylab("Predicted burst speed (mm/s)")+
  xlab("Pond")

# Showing different slopes by pond
visreg(performance_mod_wild_std, xvar = "gs", by = "pond", 
       gg = T, line = list(col = "black"), jitter = TRUE)+
  facet_wrap(~pond)+
  geom_point(alpha = 0.3)+
  ylab("Predicted burst speed (mm/s)")+
  xlab("Gosner stage")+
  theme_minimal()
```

## Tradeoff? (Dev/performance, wild)
### Make data
```{r}
## Create a new dataset with all of the main effect variables set to the mean BUT maintain treatment values.
new.wild <- wild %>% mutate(
  mass_at_stage = mean(mass_at_stage, na.rm = T),
  gs = median(gs, na.rm = T),
  PC2 = mean(PC2, na.rm = T),
  T_initial = mean(T_initial, na.rm = T),
  pond = pond,
  avg_pond_temp = avg_pond_temp
  )

# Fit the predicted speed for each individual, since we held the fixed effects constant, these are essentially just the conditional modes of the random effects. We also want to maintain our individual level variation, so we include the individual residuals from the model.
wild_tradeoff <- wild %>% 
  mutate(fit.re = predict(performance_mod_wild, newdata = new.wild, re.form = NULL),
         resid = summary(performance_mod_wild)$resid,
         fit = fit.re + resid) %>%
  select(pond, tadpole, pond, devrate, fit)
```

### Model
```{r}
# Linear model
mod <- lm(fit ~ devrate, data = wild_tradeoff)
visreg(mod)
summary(mod)
```

### Plot
```{r}
wild_tradeoff_plot <- wild_tradeoff %>%
  group_by(pond, tadpole) %>% 
  summarize(devrate = mean(devrate), fit = mean(fit)) %>%
  ggplot(aes(x = devrate, y = fit, col = pond))+
  geom_point(size = 2, alpha = 0.7)+
  geom_smooth(method = "lm", aes(col = pond, group = pond), se = F)+
  theme_minimal()+
  theme(legend.position = "bottom")+
  ylab("Predicted burst speed (mm/s)")+
  xlab("Developmental rate (GS/day)")
wild_tradeoff_plot
```

## Combined tradeoff plot (lab and wild)
```{r}


lab_tradeoff %>%
  group_by(treatment, pond, clutch, tadpole) %>%
  summarize(devrate = mean(devrate), fit = mean(fit)) %>%
  group_by(treatment, pond, clutch) %>%
  summarize(devrate = mean(devrate), fit = mean(fit)) %>%
  group_by(treatment, pond) %>%
  summarize(devrate = mean(devrate), fit = mean(fit)) %>%
  ggplot(aes(x = devrate, y = fit, col = treatment))+
  geom_point(size = 3, alpha = 0.7)+
  geom_line(aes(group = pond), lty = "dashed", lwd = 0.5, col = "black")+
  scale_color_manual(name = "Temperature treatment", values = c("orange", "dodgerblue3"))+
  geom_point(data = wild_tradeoff %>% 
               group_by(pond, tadpole) %>%
               summarize(devrate = mean(devrate), fit = mean(fit)) %>%
               group_by(pond) %>%
               summarize(devrate = mean(devrate), fit = mean(fit)),
             aes(x = devrate, y = fit),
             col = "black", size = 2, alpha = 0.9)+
  geom_smooth(data = wild_tradeoff %>%
                group_by(pond, tadpole) %>%
                summarize(devrate = mean(devrate), fit = mean(fit)),
              method = "lm", se = F, col = "black")+
  theme_minimal()+
  theme(legend.position = "bottom")+
  ylab("Predicted burst speed (mm/s)")+
  xlab("Developmental rate (GS/day)")+
  guides(col = guide_legend(override.aes = list(alpha = 1)))
# Do we want to change this so it shows the slope after accounting for the pond random effect for wild?

# Wild: points are pond averages (avg dev rate, avg burst speed). Line is fit to individual points (not shown), not the averages. Line does not currently take into account the pond random effects.
```

# Lab vs. Wild
## Allocation
Allocation rates are different between lab and wild.
```{r}
load("data/outputs/dat3_nonas.Rda")
load("data/outputs/wild_nonas.Rda")
plot.alloc <- dat3_nonas %>% ggplot(aes(x = jitter(gs, amount = 0.1), y = mass, col = treatment))+ 
  geom_point(alpha = 0.5)+
  geom_smooth(method = "lm")+
  geom_point(data = wild_nonas, color = "black", alpha = 0.5)+ # add points and line for wild tadpoles
  geom_smooth(data = wild_nonas, method = "lm", col = "black")+
  scale_color_manual(name = "Temperature treatment", values = c("orange", "dodgerblue3"))+
  theme_minimal()+
  theme(legend.position = "bottom")+
  ylab("Mass (g)")+
  xlab("Gosner stage")
plot.alloc
```

# Pond temp
Do developmental rates in the wild vary with pond temp? (mixed model: dev rates ~ 2019 pond temp + (1| pond)
```{r}
# Does developmental rate vary with pond temperature?
dev_temp_wild <- lmer(devrate ~ avg_pond_temp + (1 | pond), data = wild %>% group_by(pond, tadpole) %>% summarize(devrate = mean(devrate), avg_pond_temp = mean(avg_pond_temp)))
summary(dev_temp_wild)
visreg(dev_temp_wild, gg = T, jitter = T, line = list(col = "black"))+
  geom_point(size = 1.5, alpha = 0.5)+
  theme_minimal()+
  ylab("Predicted development rate (GS/d)")+
  xlab("2019 mean daily pond temperature (ºC)")
```
