---
title: "Morphometrics"
author: "Kaija Gahm"
date: "11/21/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup
## Load libraries
```{r, echo = FALSE}
source("libraries.R")
```

## Load data
```{r}
load("data/outputs/burst_stats.Rda")
read.csv("data/inputs/Combined_2019RespData_20190729.csv") # manual morphometrics
tpsfiles <- list.files("2019_PostResp_photos/Exposure_Edit/for_landmarks/", pattern = "*[[:digit:]].tps", full.names = T)
tpslist <- lapply(tpsfiles, readland.tps, specID = "imageID") # geometric landmarks
```

## Clean data
### Bind into one array and change dimnames
```{r}
all_landmarks <- abind(tpslist) # bind all sets of landmarks into one array

tadpoles_full <- dimnames(all_landmarks)[[3]]

tadpoles <- str_extract(tadpoles_full, pattern = "(?<=straightened\\_).*(?=\\_CH)")
tadpoles <- gsub("WIld", "Wild", tadpoles)
dimnames(all_landmarks)[[3]] <- tadpoles
```

### Fix flipped specimens
```{r}
str(all_landmarks)

# Function to flip coordinates left/right
flip_lr <- function(object){
  object[,1] <- object[,1]*-1
  return(object)
}

for(i in 1:dim(all_landmarks)[3]){
  spec <- all_landmarks[,,i]
  if(spec[1,1] > spec[11,1]){ # if the nose comes after the tail...
    print("flipped one")
    all_landmarks[,,i] <- fliplr(spec)
  }else{
    all_landmarks[,,i] <- spec
  }
}

# that looks like a reasonable number of photos to have flipped!
```



### Check for missing data
```{r}
any(is.na(all_landmarks)) # NICE
```
### Remove duplicates
```{r}
# Some of the duplicates are the three tadpoles that we deliberately duplicated to quantify inter-observer error. Others are tadpoles that we just had multiple photos of for whatever reason.
dups <- dimnames(all_landmarks)[[3]][duplicated(dimnames(all_landmarks)[[3]])]  
whichdups <- which(duplicated(dimnames(all_landmarks)[[3]]))
landmarks <- all_landmarks[,,-whichdups] # remove duplicates
```

### Make data table
```{r}
tadpoles <- dimnames(landmarks)[[3]]
ponds <- str_extract(tadpoles, pattern = "(?<=^).{2,3}(?=\\_)") %>% as.factor()
treatments <- str_extract(tadpoles, pattern = "(?<=\\_)[[:alpha:]]*(?=$)") %>% as.factor()
```

### Generate sliders
```{r}
#slidermat <- define.sliders(landmarks = landmarks, nsliders = 15)
slidermat <- data.frame(before = c(7, 12, 13, 14, 7, 20, 21, 22, 9, 24, 25, 8, 16, 17, 18),
                        slide = c(12, 13, 14, 15, 20, 21, 22, 23, 24, 25, 26, 16, 17, 18, 19),
                        after = c(13, 14, 15, 11, 21, 22, 23, 11, 25, 26, 11, 17, 18, 19, 11)) %>% as.matrix()
```

## Analysis
### Generalized Procrustes Analysis
```{r}
tadpole.gpa <- gpagen(A = landmarks, curves = slidermat)
plot(tadpole.gpa)
```

### Plot PCA
```{r}
# conduct PCA
pca <- plotTangentSpace(tadpole.gpa$coords, groups = treatments) # note that `treatments` is a separate vector, not a part of tadpole.gpa.
str(pca, 1)

# separate out the scores data
scores <- pca$pc.scores %>% as.data.frame() %>% mutate(tadpole = tadpoles, treatment = treatments, pond = ponds)

# separate out the importances data
pca.importance <- pca$pc.summary$importance %>% t() %>% as.data.frame() %>% 
  mutate(pcname = row.names(.),
         pcnum = 1:nrow(.)) %>% 
  dplyr::rename(stdev = "Standard deviation", propvar = "Proportion of Variance", cumprop = "Cumulative Proportion")

# PC scores plot
scores %>% ggplot(aes(x = PC1, y = PC2, col = treatment))+
  geom_point() + ggtitle("Tadpole morphometrics score plot") + stat_ellipse()

# PC importance
pca.importance %>% ggplot(aes(x = pcnum, y = cumprop)) +
  geom_col() +
  ggtitle("Cumulative variation explained by principal components")+
  ylab("Cumulative proportion of variation")+
  xlab("PC number")

pca.importance %>% ggplot(aes(x = pcnum, y = propvar)) +
  geom_col() +
  ggtitle("Variation explained by each principal component")+
  ylab("Proportion of variation explained")+
  xlab("PC number")

```

#### Investigate PCA outliers
```{r}
scores <- scores %>% 
  mutate(investigate = case_when(PC1 < -0.15 ~ T,
                                 TRUE ~ F))
outliers <- scores %>% filter(investigate == T) %>% pull(tadpole)

# were any of these duplicates?

outliers %in% dups

# Search through manually to find them (ugh, there has to be a better way)
outliers

# B0_02_B_Low had misaligned landmarks
#BS_03_A_High is just a tiny little squirt--exclude?
# LO_04_C_High looks normal... hmm.
```

### Visualize individuals vs. mean
```{r}
ref <- mshape(tadpole.gpa$coords) # mean coordinate values

res <- plotTangentSpace(tadpole.gpa$coords, groups = treatments, verbose=TRUE)

# Plot extremes of PC1
parameters_PC1 <- gridPar(tar.pt.bg = "red", pt.size = 1, n.col.cell=50)

plotRefToTarget(M1 = ref, M2 = res$pc.shapes$PC1min, gridPars = parameters_PC1, method = "TPS") #shape change along PC1 in the negative direction

plotRefToTarget(M1 = ref, M2 = res$pc.shapes$PC1max, gridPars = parameters_PC1, method = "TPS") #shape change along PC1 in the positive direction


# Plot extremes of PC2
parameters_PC2 <- gridPar(tar.pt.bg = "blue", pt.size = 1, n.col.cell=50)

plotRefToTarget(M1 = ref, M2 = res$pc.shapes$PC2min, gridPars = parameters_PC2, method = "TPS") #shape change along PC1 in the negative direction

plotRefToTarget(M1 = ref, M2 = res$pc.shapes$PC2max, gridPars = parameters_PC2, method = "TPS") #shape change along PC1 in the positive direction
```
 
So it seems that PC1 is basically a measurement of relative head/tail lengths, and PC2 is basically a measurement of length/depth ratio. Cool! Much more informative than linear measurements would have been.


## Investigate error
### Prepare data
```{r}
# Back to the original dataset, before we removed the duplicates
str(all_landmarks)

repeats_1 <- which(duplicated(dimnames(all_landmarks)[[3]], fromLast = F))
repeats_2 <- which(duplicated(dimnames(all_landmarks)[[3]], fromLast = T))
allrepeats <- union(repeats_1, repeats_2)
  
repeats <- all_landmarks[,,allrepeats] # select the repeated specimens to use

# Now, we only want the ones that were deliberate repeats of individual photographs, rather than different photos of the same specimen.
freqrepeats <- table(dimnames(repeats)[[3]]) %>% as.data.frame()

# subset out the deliberately repeated photos
deliberate_repeats_names <- freqrepeats$Var1[freqrepeats$Freq == 20]

deliberate_repeats_list <- vector(mode = "list", length = 3) #empty list

for(i in 1:3){
  name <- deliberate_repeats_names[i]
  deliberate_repeats_list[[i]] <- all_landmarks[,,which(dimnames(all_landmarks)[[3]] == name)]
}
```

### Get errors
```{r}
# Procrustes analysis on each group of landmarks
gpa_1 <- gpagen(A = deliberate_repeats_list[[1]], curves = slidermat)
plot(gpa_1)
gpa_2 <- gpagen(A = deliberate_repeats_list[[2]], curves = slidermat)
plot(gpa_2)
gpa_3 <- gpagen(A = deliberate_repeats_list[[3]], curves = slidermat)
plot(gpa_3)

# Get variances
var_1 <- gpa_1$points.var %>% as.data.frame() %>% mutate(sample = "A", landmark = row.names(.))
var_2 <- gpa_2$points.var %>% as.data.frame() %>% mutate(sample = "B", landmark = row.names(.))
var_3 <- gpa_3$points.var %>% as.data.frame() %>% mutate(sample = "C", landmark = row.names(.))

vars <- rbind(var_1, var_2, var_3)
vars2 <- vars %>% group_by(as.numeric(landmark)) %>% 
  summarize(x = mean(Var.X),
            y = mean(Var.Y)) %>% 
  rename(landmark = `as.numeric(landmark)`) %>% pivot_longer(cols = c(x, y), names_to = "dimension", values_to = "variance") %>% mutate(label = "within") %>% as.data.frame()

# Plot variances
vars2 %>% ggplot(aes(x = landmark, y = variance))+
  geom_point(aes(col = dimension)) +
  ggtitle("Measurement variance in test photos")+
  ylab("Variance (cm)") +
  xlab("Landmark")
```

### Compare to variances across photos
```{r}
# Get between-specimen variances
allvar <- tadpole.gpa$points.var %>% as.data.frame() %>% 
  mutate(landmark = 1:nrow(.)) %>% rename(x = Var.X, y = Var.Y) %>%
  pivot_longer(cols = c(x, y), names_to = "dimension", values_to = "variance") %>% mutate(label = "between")

# Join with within-specimen errors
between_within <- rbind(vars2, allvar)

# Plot to see differences
between_within %>% ggplot(aes(x = landmark, y = variance, col = label))+
  geom_point() +
  facet_wrap(~dimension)

# This doesn't seem good...

```


