---
title: "Step5_Edit_Bursts.Rmd"
author: "Kaija Gahm"
date: "11/21/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup
## Load libraries
```{r, echo = FALSE}
source("libraries.R")
source("tally_by_group.R")
source("process_coordinates_funs.R")
```

## Load data
```{r}
load("data/outputs/speedsDF.Rda")
load("data/outputs/trial_start_frames.Rda")
load("data/outputs/scaled_coords_df.Rda")
```

## Assign frame numbers to speedsDF and scaled_coords_df
```{r}
speedsDF <- speedsDF %>% mutate(unique_trial = paste(fullname, trial)) %>% as.data.frame()
speedsDF <- tally_by_group(speedsDF, group_col = "unique_trial")

scaled_coords_df <- scaled_coords_df %>% mutate(unique_trial = paste(fullname, trial)) %>% as.data.frame()
scaled_coords_df <- tally_by_group(scaled_coords_df, group_col = "unique_trial")
```

## Trim trial beginnings
### Join trials to start frames
```{r}
trial_start_frames <- trial_start_frames %>% mutate(unique_trial = paste(fullname, trial))

speedsDF <- left_join(speedsDF, trial_start_frames, by = c("fullname", "trial", "unique_trial"))
scaled_coords_df <- left_join(scaled_coords_df, trial_start_frames, by = c("fullname", "trial", "unique_trial"))
```

### Add initial (stationary) frame
```{r}
# We want to allow one extra frame before the start of the trial for padding, so we can set the speed to 0 right before the trial starts.
speedsDF <- speedsDF %>% mutate(frame_init = frame_start - 1)

scaled_coords_df <- scaled_coords_df %>% mutate(frame_init = frame_start - 1)

# Set speed to 0 for the initial frame (speedsDF)
speedsDF$speedma5[speedsDF$frame == speedsDF$frame_init] <- 0
speedsDF$speed[speedsDF$frame == speedsDF$frame_init] <- 0

# Set displacement to 0 for the initial frame (scaled_coords_df)
scaled_coords_df$displacement[scaled_coords_df$frame == scaled_coords_df$frame_init] <- 0
scaled_coords_df$displacement[scaled_coords_df$frame == scaled_coords_df$frame_init] <- 0
```

### Remove frames before initial frame
```{r}
speedsDF <- speedsDF %>% filter(frame >= frame_init)
scaled_coords_df <- scaled_coords_df %>% filter(frame >= frame_init)
```

### Relative vs. absolute frames
```{r}
speedsDF <- speedsDF %>% rename(frame_trial = frame)
speedsDF <- tally_by_group(speedsDF, group_col = "unique_trial") %>% rename(frame_burst = frame)

scaled_coords_df <- scaled_coords_df %>% rename(frame_trial = frame)
scaled_coords_df <- tally_by_group(scaled_coords_df, group_col = "unique_trial") %>% rename(frame_burst = frame)
```

## Duration of bursts
### Check max durations
```{r}
tooshort <- speedsDF %>% group_by(fullname, trial) %>% 
  summarize(maxframe = max(frame_burst)) %>%
  filter(maxframe < 60)

nrow(tooshort)
# there are 74 bursts that are less than 60 frames long. What about less than half a sec?

waytooshort <- speedsDF %>% group_by(fullname, trial) %>% 
  summarize(maxframe = max(frame_burst)) %>%
  filter(maxframe < 30)
nrow(waytooshort)

# Only 14 are less than half a second. Maybe we should exclude those?
```

Need to discuss whether to exclude bursts less than 1/2 sec.
```{r}
#speedsDF <- speedsDF %>% filter(!(paste(fullname, trial) %in% paste(waytooshort$fullname, waytooshort$trial))) # Remove individual bursts that were "waytooshort".
```


## Trim burst lengths
### Trim speeds
```{r}
todrop <- c("time", "angle", "px", "date", "frame_trial", "frame_start", "unique_trial")

bursts <- speedsDF %>% select(-todrop) %>% mutate(crop = "Full burst")
bursts_sec <- bursts %>% filter(frame_burst <= 60) %>% mutate(crop = "First second")
bursts_halfsec <- bursts %>% filter(frame_burst <= 30) %>% mutate(crop = "First half second")
bursts_secondhalfsec <- bursts %>% filter(frame_burst <= 60 & frame_burst > 30) %>% mutate(crop = "Second half second")
bursts_allcrops <- rbind(bursts, bursts_sec, bursts_halfsec, bursts_secondhalfsec)
```

### Trim scaled coords
```{r}
todrop <- c("time", "angle", "px", "date", "frame_trial", "frame_start", "unique_trial")

scoords <- scaled_coords_df %>% select(-todrop) %>% mutate(crop = "Full burst")
scoords_sec <- scoords %>% filter(frame_burst <= 60) %>% mutate(crop = "First second")
scoords_halfsec <- scoords %>% filter(frame_burst <= 30) %>% mutate(crop = "First half second")
scoords_secondhalfsec <- scoords %>% filter(frame_burst <= 60 & frame_burst > 30) %>% mutate(crop = "Second half second")
scoords_allcrops <- rbind(scoords, scoords_sec, scoords_halfsec, scoords_secondhalfsec)
```

## Fix NA's in speeds
```{r}
# Look at where the NA's in speed are
bursts_allcrops[is.na(bursts_allcrops$speed),] #All of them have frame_burst as 1, so let's set the speed to 0. 

# Set speed to 0
bursts_allcrops$speed[is.na(bursts_allcrops$speed)] <- 0
```

## Add try numbers and edit full names
```{r}
trynums <- str_extract(string = bursts_allcrops$fullname, pattern = "(?<=\\_)[[:digit:]]{1}(?=\\_2019)") # get try numbers

bursts_allcrops$trynum <- trynums # edit try numbers
scoords_allcrops$trynum <- trynums # edit try numbers

# Set only tries to 1
bursts_allcrops$trynum[is.na(bursts_allcrops$trynum)] <- "1"
scoords_allcrops$trynum[is.na(scoords_allcrops$trynum)] <- "1"
```

## Edit full names to remove try numbers
```{r}
newfullnames <- str_replace(string = bursts_allcrops$fullname, pattern = "\\_[[:digit:]]{1}(?=\\_2019)", replacement = "") # get new full names (note pattern includes the _)

bursts_allcrops$fullname <- newfullnames # edit full names
scoords_allcrops$fullname <- newfullnames # edit full names
```

## Add tadpole names
```{r}
bursts_allcrops <- bursts_allcrops %>% mutate(tadpole = str_extract(fullname, pattern = "(?<=^).*(?=\\_[[:digit:]]{8}\\_.*$)"))
# The above line doesn't quite work to get the names for the wild tadpoles, so we have to add "Wild" on.
bursts_allcrops$tadpole[bursts_allcrops$treatment == "Wild"] <- paste0(bursts_allcrops$tadpole[bursts_allcrops$treatment == "Wild"], "_Wild")

scoords_allcrops <- scoords_allcrops %>% mutate(tadpole = str_extract(fullname, pattern = "(?<=^).*(?=\\_[[:digit:]]{8}\\_.*$)"))
# The above line doesn't quite work to get the names for the wild tadpoles, so we have to add "Wild" on.
scoords_allcrops$tadpole[scoords_allcrops$treatment == "Wild"] <- paste0(scoords_allcrops$tadpole[scoords_allcrops$treatment == "Wild"], "_Wild")
```

## Double check that we're not missing any multi-try tadpoles
Try numbers were derived from the tadpoles' full names, which were in turn derived from the file names of the photos. I don't trust those, so let's double check.
```{r}
bursts_allcrops <- bursts_allcrops %>% mutate(fulldate = str_extract(fullname, "[[:digit:]]{8}"))

tries <- bursts_allcrops %>% group_by(tadpole) %>% mutate(ntries = length(unique(fulldate))) %>% group_by(tadpole, fulldate) %>% slice(1) %>% as.data.frame()

tries_numbered <- tries %>% tally_by_group(group_col = "tadpole") %>% rename(trynum_new = frame) %>% select(fullname, trynum_new)

bursts_allcrops <- bursts_allcrops %>% left_join(tries_numbered, by = "fullname") %>% mutate(trynum = NULL) %>% rename(trynum = trynum_new) # got rid of the old try numbers and added the new ones.


#~~~~~~~~~~~~~~~
# Same for scoords
scoords_allcrops <- scoords_allcrops %>% mutate(fulldate = str_extract(fullname, "[[:digit:]]{8}"))

tries <- scoords_allcrops %>% group_by(tadpole) %>% mutate(ntries = length(unique(fulldate))) %>% group_by(tadpole, fulldate) %>% slice(1) %>% as.data.frame()

tries_numbered <- tries %>% tally_by_group(group_col = "tadpole") %>% rename(trynum_new = frame) %>% select(fullname, trynum_new)

scoords_allcrops <- scoords_allcrops %>% left_join(tries_numbered, by = "fullname") %>% mutate(trynum = NULL) %>% rename(trynum = trynum_new) # got rid of the old try numbers and added the new ones.
```

## Remove first trynums for tadpoles with multiple tries
Biological reasoning: If I did a second try, it was because the first one(s) were inadequate/bad, so we will exclude those.
```{r}
bursts_allcrops <- bursts_allcrops %>% group_by(tadpole) %>% filter(trynum == max(trynum)) %>% as.data.frame()

# Double check that we only have one try per tadpole.
test <- bursts_allcrops %>% group_by(tadpole) %>% summarize(ntries = length(unique(trynum)))
table(test$ntries, exclude = NULL)

# ~~~~~~~~
# now the same for scoords
scoords_allcrops <- scoords_allcrops %>% group_by(tadpole) %>% filter(trynum == max(trynum)) %>% as.data.frame()

# Double check that we only have one try per tadpole.
test <- scoords_allcrops %>% group_by(tadpole) %>% summarize(ntries = length(unique(trynum)))
table(test$ntries, exclude = NULL)
```

## Output
```{r}
save(bursts_allcrops, file = "data/outputs/bursts_allcrops.Rda")
save(scoords_allcrops, file = "data/outputs/scoords_allcrops.Rda")
```


