---
title: 'Step 4: Find Burst Beginnings'
author: "Kaija Gahm"
date: "10/4/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup
## Load libraries
```{r, echo = FALSE}
source("libraries.R")
source("tally_by_group.R")
source("process_coordinates_funs.R")
```

## Load data
```{r}
load("data/outputs/speedsDF.Rda")
```

# Pixels and angles
## Standardize angles
```{r}
speedsDF <- speedsDF %>% group_by(fullname, trial) %>%
  mutate(anglet = angleTransform(angle))
```

## Angle moving averages
```{r}
speedsDF <- speedsDF %>% group_by(fullname, trial) %>%
  mutate(atma5 = rollapplyr(anglet, 5, mean, fill = NA))
```

## Pixel moving averages
```{r}
speedsDF <- speedsDF %>% group_by(fullname, trial) %>%
  mutate(pxma5 = rollapplyr(px, 5, mean, fill = NA))
```

# Training
For some trials, I went in manually and identified the starting frame. Now I'm going to try to find a way to automatically identify the starting frame, using this data as a means for testing any proposed models/methods.

## Load training data
```{r}
trueframes <- read.csv("data/inputs/getframenums.csv")
trueframes <- trueframes[,c(1, 5)]
names(trueframes) <- c("fullname", "frame")
trueframes$fullname <- gsub("_edit|.avi", "", trueframes$fullname)
trueframes$trial <- 1 # all the testing data was using trial 1, because those were the easiest trials to do by hand, and because we have no expectation that different trials should behave differently.
head(trueframes)
```

## Beginnings (three methods)
```{r}
trueframes <- trueframes[trueframes$fullname %in% speedsDF$fullname,]
for(i in 1:nrow(trueframes)){
    df <- speedsDF %>% filter(fullname == trueframes$fullname[i], trial == "1") # pick out the video and trial
    trueframes$beg_sp[i] <- findBeginningSpeed(df, "speedma5") # find beginning using speed
    trueframes$beg_ag[i] <- findBeginningAngle(df, "atma5") # find beginning using angle
    trueframes$beg_px[i] <- findBeginningPx(df, "pxma5") # find beginning using pixels
}
```

## Residuals (three methods)
```{r}
trueframes$spdiff <- trueframes$beg_sp - trueframes$frame # speed residual
trueframes$agdiff <- trueframes$beg_ag - trueframes$frame # angle residual
trueframes$pxdiff <- trueframes$beg_px - trueframes$frame # pixel residual
```

## Z-scores (three methods)
```{r}
# Calculate z-scores for the differences. I still don't quite get this.
trueframes$spz <- (trueframes$spdiff - mean(trueframes$spdiff))/std(trueframes$spdiff)
trueframes$agz <- (trueframes$agdiff - mean(trueframes$agdiff))/std(trueframes$agdiff)
trueframes$pxz <- (trueframes$pxdiff - mean(trueframes$pxdiff))/std(trueframes$pxdiff)

# Visualize z-scores for each method
ggplot(trueframes, aes(x = fullname))+
  geom_point(aes(y = spz), col = "blue")+
  geom_point(aes(y = agz), col = "red", alpha = 0.9)+
  geom_point(aes(y = pxz), col = "green", alpha = 0.7) +
  ylab("Z-score")
```

## Individual models (three methods)
```{r}
# Make individual models for each method
spm <- lm(frame~beg_sp, data = trueframes)
pxm <- lm(frame~beg_px, data = trueframes)
agm <- lm(frame~beg_ag, data = trueframes)
# View the model results
summary(spm)
summary(pxm)
summary(agm)
```

## Combined models
```{r}
allm <- lm(frame~beg_sp+beg_ag+beg_px, data = trueframes)
summary(allm)
# notice that the sp predictor isn't significant; what if we remove it?
```

```{r}
agpxm <- lm(frame~beg_ag+beg_px, data = trueframes)
summary(agpxm)
plot(fitted(agpxm)~trueframes$frame)
trueframes$fitted <- fitted(agpxm)
trueframes$resid <- trueframes$fitted-trueframes$frame
# This model does a good job of predicting the frame, assuming we will have removed outliers first. 
```

# How will we separate "by hand" trials--outliers
## Variance in estimates
```{r}
trueframes$estimates_var <- rowVars(as.matrix(trueframes[,c("beg_sp", "beg_ag", "beg_px")]), na.rm = T)
```

## Find true problems
### Using combined model
```{r}
trueframes$problem <- ifelse(abs(trueframes$resid) > 3, T, F)
table(trueframes$problem)
# Yikes, that's bad. If trueframes$problem == TRUE, that means that we failed to identify a trial whose predicted start frame was way off from the true (by hand) start frame.
```

### Using angles only
```{r}
trueframes$problem <- ifelse(abs(trueframes$agdiff) > 3, T, F) # For testing purposes, identify the ones in this dataset that are really wrong, based on angle
table(trueframes$problem)
# That's better: leave it like this. Fewer TRUE's is good because that means we failed in fewer cases.
```

## Test variance thresholds
### Thresholds
```{r}
# Make a data frame of thresholds to test
threshes <- data.frame(
  thresh = seq(1, 30, by = 0.1),
  fnr = NA,
  fpr = NA,
  prop_byhand = NA,
  total_fn_prop = NA
)
```

### Test thresholds
```{r}
# Calculate false negative and false positive rates for each threshold
for(i in 1:nrow(threshes)){
  namesflagged <- trueframes$fullname[trueframes$estimates_var > threshes$thresh[i]]
  namesnotflagged <- trueframes$fullname[trueframes$estimates_var <= threshes$thresh[i]]
  namespos <- trueframes$fullname[trueframes$problem == T]
  namesneg <- trueframes$fullname[trueframes$problem == F]
  npos = length(namespos)
  nneg = length(namesneg)
  nflagged <- length(namesflagged)
  nnotflagged <- length(namesnotflagged)

  names_fn <- namespos[namespos %in% namesnotflagged]
  names_fp <- namesneg[namesneg %in% namesflagged]
  nfn <- length(names_fn) # number of false negatives 
  nfp <- length(names_fp) # number of false positives
  
  fnr <- nfn/npos # false negative rate
  threshes$fnr[i] <- fnr
  total_fn_prop <- nfn/99
  
  threshes$total_fn_prop[i] <- total_fn_prop # proportion of the total data that will be false negatives
  
  fpr <- nfp/nneg # false positive rate
  threshes$fpr[i] <- fpr
  
  prop_byhand = nflagged/99 #what proportion of the data will we have to do by hand
  threshes$prop_byhand[i] <- prop_byhand
}
```

### See results
```{r}
# Convert threshes to a long data frame for plotting
threshes_long <- melt(threshes, id.vars = "thresh", variable.name = "rate")

# Plot thresholds vs. how many it flags
ggplot(threshes_long, aes(x = thresh, y = value, col = rate))+
  geom_line()+
  xlab("Variance threshold")+
  ylab("Proportion")
```

### Is there any way we can reduce the false negative rate?
```{r}
# find the false negatives in this test data set
trueframes$byhand <- ifelse(trueframes$estimates_var > 5, T, F)
fns <- trueframes[trueframes$byhand == F & trueframes$problem == T,]
fns_names <- fns$fullname

#nothing obvious is going on, which is too bad, bc that means we can't filter it out
summary(fns$agdiff) # how different is the angle estimate from the true frame? Not too bad.
summary(fns$resid) # If we use the residuals from the combined model, it won't be bad at all.
```

## Which would be by hand?
```{r}
trueframes$byhand <- ifelse(trueframes$estimates_var > 5, T, F) # which ones will we have to do by hand
```

# Apply protocol
```{r}
head(speedsDF)
#  how many unique trials?
length(unique(speedsDF[,c("fullname", "trial")])$fullname)
```

## Get unique trials
```{r}
# Find all the unique trials
starts <- data.frame(
  fullname = unique(speedsDF[,c("fullname", "trial")])$fullname,
  trial = unique(speedsDF[,c("fullname", "trial")])$trial
)
```

## [Compute beginnings] VERY SLOW--SKIP TO NEXT CODE CHUNK TO LOAD PRE-RUN DATA
```{r, eval = F}
for(i in 1:nrow(starts)){
  tryCatch(
    expr = {
      df <- speedsDF %>% filter(fullname == starts$fullname[i], trial == starts$trial[i])
      starts$beg_sp[i] <- findBeginningSpeed(df, "speedma5")
      starts$beg_ag[i] <- findBeginningAngle(df, "atma5")
      starts$beg_px[i] <- findBeginningPx(df, "pxma5") + 1
      message("Iteration ", i, " successful.")
    },
    error = function(e){
      message("* Caught an error on iteration", i)
      print(e)
    }
  )
}
save(starts, file = "data/outputs/starts.Rda")
```

## Load pre-computed beginnings
```{r}
#save(starts, file = "starts.Rda")
load("data/outputs/starts.Rda")
```

## Prepare data
### Replace Inf with NA
```{r}
# Function
inf2NA <- function(vec){
  vec[is.infinite(vec)] <- NA
  return(vec)
}

# Apply the function
starts2 <- starts %>% mutate_each(funs(inf2NA))
```

### Compute estimates variance
```{r}
starts2$estimates_var <- rowVars(as.matrix(starts2[,c("beg_sp", "beg_ag", "beg_px")]), na.rm = T)
```

## Separate automatic/manual
```{r}
# Flag observations with var > 5
starts2$byhand <- ifelse(starts2$estimates_var >= 5, T, F)
# What percent of the observations were flagged as "by hand"?
(pct_byhand <- sum(starts2$byhand, na.rm = T)/nrow(starts2))

# Which ones do we need to do by hand
todobyhand <- starts2 %>% filter(byhand == TRUE)

# Which ones can we do automatically
automatics <- starts2 %>% filter(byhand == FALSE)
```

# Get beginnings
## Automatic
### Compute using model
```{r}
automatics <- automatics %>% mutate(fitted = predict.lm(agpxm, newdata = automatics)) %>% mutate_at(vars(fitted), funs(round))
```

### Subset automatic data
```{r}
automatics_1 <- automatics %>% rename(frame_start = fitted) %>% select(fullname, trial, frame_start) %>% mutate(method = "automatic")
```


## Manual
### Save data
```{r}
#write.csv(todobyhand, file = paste("data/outputs/", Sys.time(), "todobyhand.csv"))
```

### Load manual data
```{r}
manuals <- read.csv("data/inputs/manual_kaija.csv", header = T, as.is = T, na.strings = c("NA", ""))
```

### Clean/subset manual data
```{r}
# Calculate starting frame with reference to the beginning of the _trial_
manuals <- manuals %>% mutate_at(vars(verification), list(as.factor))
manuals_1 <- manuals %>% rename(frame_start = difference) %>% select(fullname, trial, frame_start) %>% mutate(method = "manual")
```

## Merge automatic and manual data
```{r}
trial_start_frames <- rbind(automatics_1, manuals_1) %>% as.data.frame()

# order by fullname and trial
trial_start_frames <- trial_start_frames %>% arrange(fullname, trial)

```

# Output
```{r}
save(trial_start_frames, file = "data/outputs/trial_start_frames.Rda")
```


