---
title: "Step5_Edit_Bursts.Rmd"
author: "Kaija Gahm"
date: "11/21/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup
## Load libraries
```{r, echo = FALSE}
source("libraries.R")
source("tally_by_group.R")
source("process_coordinates_funs.R")
```

## Load data
```{r}
load("data/outputs/speedsDF.Rda")
load("data/outputs/trial_start_frames.Rda")
```

## Assign frame numbers to speedsDF
```{r}
speedsDF <- speedsDF %>% mutate(unique_trial = paste(fullname, trial)) %>% as.data.frame()
speedsDF <- tally_by_group(speedsDF, group_col = "unique_trial")
```

## Trim trial beginnings
### Join trials to start frames
```{r}
trial_start_frames <- trial_start_frames %>% mutate(unique_trial = paste(fullname, trial))
speedsDF <- left_join(speedsDF, trial_start_frames, by = c("fullname", "trial", "unique_trial"))
```

### Remove frames < start frame
```{r}
speedsDF <- speedsDF %>% filter(frame >= frame_start)
```

### Relative vs. absolute frames
```{r}
speedsDF <- speedsDF %>% rename(frame_trial = frame)
speedsDF <- tally_by_group(speedsDF, group_col = "unique_trial") %>% rename(frame_burst = frame)
```

## Trim burst lengths
```{r}
todrop <- c("time", "angle", "px", "date", "frame_trial", "frame_start", "unique_trial")

bursts <- speedsDF %>% select(-todrop) %>% mutate(crop = "Full burst")
bursts_sec <- bursts %>% filter(frame_burst <= 60) %>% mutate(crop = "First second")
bursts_halfsec <- bursts %>% filter(frame_burst <= 30) %>% mutate(crop = "First half second")
bursts_allcrops <- rbind(bursts, bursts_sec, bursts_halfsec)
```

## Output
```{r}
save(bursts_allcrops, file = "data/outputs/bursts_allcrops.Rda")
```


